### 2023-10-10 niceAi.Rmd
initial set up using historical HMM modelling

---
title: "Untitled"
author: "Ismail Guennouni"
date: "2023-05-16"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(depmixS4)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
origin_dat <- read.csv("data/longFormData.csv")

origin_dat <- origin_dat %>% mutate(investment_bin = cut(investment, breaks = c(-1,2.5,7.5,12.5,17.5,21)),
                                    return_pct_bin = cut(return_pct, breaks = c(-.1,.125,.325,.51,.755,1.1)),
                                    next_investment = lead(investment, default=0),
                                            invPnL = return - investment,
                                            trPnL = 3*investment - return)
```

# Support functions and discrete Gaussian class construction 
```{r supportHMM, include=FALSE}
# order the states of the HMM to allow it to order them as low ret/mid/ret/high ret. 
label_switch <- function(mod,labels) {
  # labels is vector, first element is new label for original state 1, second is new label for original state 2, etc.
  if(!is(mod,"depmix") || !is(mod,"depmix.fitted")) stop("this function is for depmix models")
  n_states <- mod@nstates
  if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
    stop("labels needs to be a vector of unique integers between 1 and", n_states)
  }
  inv_labels <- sapply(1:n_states,function(x) which(labels == x))
  tmp <- mod
  # relabel prior
  ppars <- getpars(mod@prior)
  fpars <- getpars(mod@prior,which="fixed")
  out_pars <- as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels]))
  out_fixed <- as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels]))
  if(!tmp@prior@family$link=="identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
  # relabel transition
  for(i in 1:n_states) {
    ppars <- getpars(mod@transition[[inv_labels[i]]])
    fpars <- getpars(mod@transition[[inv_labels[i]]], which="fixed")
    out_pars <- c(out_pars,as.numeric(t(matrix(ppars,nrow=length(ppars)/n_states,byrow = TRUE)[,inv_labels])))
    out_fixed <- c(out_fixed,as.logical(t(matrix(fpars,nrow=length(fpars)/n_states,byrow = TRUE)[,inv_labels])))
    tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
    if(!tmp@transition[[i]]@family$link=="identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
    #out_pars <- c(out_pars,getpars(mod@transition[[inv_labels[i]]]))
  }
  # relabel response
  for(i in 1:n_states) {
    out_pars <- c(out_pars,unlist(lapply(mod@response[[inv_labels[i]]],getpars)))
    out_fixed <- c(out_fixed,unlist(lapply(mod@response[[inv_labels[i]]],getpars,which="fixed")))
  }
  tmp <- setpars(tmp,out_fixed,which="fixed")
  tmp <- setpars(tmp,out_pars)
  if(is(tmp,"depmix.fitted")) tmp@posterior <- viterbi(tmp)
  return(tmp)
}

# ORDERING STATES WHEN RESPONSE IS BINS
order_mod <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation (5 bins, for binned returns)
    sum[i] = sum(1:5*mod@response[[i]][[1]]@parameters$coefficients)
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum))
  return(mod)
}

# ORDERING STATES WHEN RESPONSE IS GAUSSIAN
order_mod_gauss <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0,ns)
  for(i in 1:ns) {
    # Expected return in each state calculation 
    sum[i] = mod@response[[i]][[1]]@parameters$mu
  }
  # reordering the states
  mod <- label_switch(mod, rank(sum))
  return(mod)
}

```

```{r, include=F}
# define a response class which only contains the standard slots, no additional slots
setClass("discgaus", contains="response", slots=c(breaks="numeric"))

# define a generic for the method defining the response class

setGeneric("discgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("discgaus"))

# define the method that creates the response class

setMethod("discgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("discgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks)
            mod
          }
)

setMethod("show","discgaus",
          function(object) {
            cat("Gaussian with discrete support\n")
            cat("Parameters: \n")
            cat("mu: ", object@parameters$mu, "\n")
            cat("sigma: ", object@parameters$sigma, "\n")
          }
)

setMethod("dens","discgaus",
          function(object,log=FALSE) {
            p <- pnorm(object@breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(object@breaks[-length(object@breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

setMethod("setpars","discgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values)!=npar) stop("length of 'values' must be",npar)
            # determine whether parameters or fixed constraints are being set
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

setMethod("getpars","discgaus",
          function(object,which="pars",...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

setMethod("fit","discgaus",
          function(object,w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object,c(pars[1], exp(pars[2])))
            object
          }
)

setClass("truncdiscgaus", contains="discgaus", slots=c(min="numeric", max="numeric"))

setMethod("dens","truncdiscgaus",
          function(object,log=FALSE) {
            breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
            p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            p <- p/sum(p)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

setGeneric("truncdiscgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("truncdiscgaus"))

setMethod("truncdiscgaus", 
          signature(y="ANY"), 
          function(y,pstart=NULL,fixed=NULL, breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20, ...) {
            y <- matrix(y,length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0,npar))
            if(!is.null(pstart)) {
              if(length(pstart)!=npar) stop("length of 'pstart' must be ",npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("truncdiscgaus",parameters=parameters,fixed=fixed,x=x,y=y,npar=npar, breaks=breaks, min=min, max=max)
            mod
          }
)
```

```{r}
# Specify simple models to copy transition and prior
inv_simple <- list()
for(i in 2:6) {
  inv_simple[[i]] <- depmix(investment_bin ~ 1, data = origin_dat, nstates = i, transition = ~ invPnL, family=multinomial("mlogit"), ntimes = rep(10,381))
}

# Specify truncated gaussian models for 3 state model 
rModels3 <- list(
  list(
    truncdiscgaus(origin_dat$investment,pstart=c(5,2), min=-.5, max=20.5) # first guassian initialised with mean of 5 and sd of 2
  ),
  list(
    truncdiscgaus(origin_dat$investment,pstart=c(10,2), min=-.5, max=20.5)
  ),
  list(
    truncdiscgaus(origin_dat$investment,pstart=c(15,2), min=-.5, max=20.5)
  )
)

inv_dmod_3 <- makeDepmix(response=rModels3,transition=inv_simple[[3]]@transition,prior=inv_simple[[3]]@prior,ntimes = rep(10,381), homogeneous=FALSE)

inv_fdmod_3 <- fit(inv_dmod_3, emcontrol=em.control(random.start = FALSE))  #random.start = FALSE -> use parameters supplied as starting values

mod_3S <- order_mod_gauss(inv_fdmod_3)

```

```{r, include=F}
#ORDER THE STATES OF THE HMM

state1 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[1]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))

state2 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[2]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))

state3 <- dens(truncdiscgaus(seq(0:20) - 1,pstart=c(unlist(mod_3S@response[[3]][[1]]@parameters)), breaks = c(-Inf, 0:19 + .5, Inf), min=0, max=20))

investment <- seq(0:20) - 1

inv_resp_3S <- as.data.frame(cbind(investment,state1,state2,state3)) %>% 
  pivot_longer(cols=c("state1","state2","state3"),
                    names_to='investor_state',
                    values_to='probability')

```

```{r}
inv_resp_3S %>% mutate( investor_state = fct_recode(investor_state, "low-trust" = "state1", "medium-trust" = "state2", "high-trust" = "state3"))%>%
        ggplot( aes(x = investment,
                    y = probability,
                    fill = investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state') +
  theme_bw() + 
  theme(legend.position = "bottom")

```

Let's write a function to plot transition functions given parameters

```{r}
# Here pars takes the form of a list ns matrices, one for each state from 1 to ns.
plot_transitions_PnL <- function(fmod) {

  ns <- nstates(fmod)

  trans_prob <- data.frame(
    from = rep(1:ns, each=80*ns),
    to = rep(1:ns, each=80),
    PnL = seq(-20,60,length=80),
    prob = 0
  )

  y <- matrix(0.0,ncol=ns, nrow=80)

  for(from in 1:ns) {
    pars <- matrix(getpars(fmod)[seq(ns + 1 + (from - 1)*2*ns, ns + 1 + (from - 1)*2*ns + 2*ns - 1)],ncol=2)
    for(to in 1:ns) {
      x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"PnL"]
      y[,to] <- exp(pars[to,1] + pars[to,2]*x)
    }
    y <- y/rowSums(y)
    for(to in 1:ns) {
      trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
    }
  }

  trans_prob <- trans_prob %>% mutate(from=recode(from, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust"),
                        to=recode(to, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust")) %>% 
    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))

#     iris %>%
#    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>%
# ggplot() + 
#    geom_histogram(aes(Petal.Width))+ 
#    facet_grid(Species~.)
#   
  ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1)) +
      theme(legend.position = "bottom")
}
```

```{r}
plot_transitions_PnL(mod_3S)
```
```{r}
ns <- 3
pars <- NULL

for (from in 1:ns){
  Tr_mat <- matrix(getpars(mod_3S)[seq(ns + 1 + (from - 1)*2*ns, ns + 1 + (from - 1)*2*ns + 2*ns - 1)],ncol=2)

colnames(Tr_mat) <- c("intercept","invPnL")
rownames(Tr_mat) <- c("to_1", "to_2", "to_3")
pars[[from]] <- Tr_mat

}

pars[[2]]["to_2","invPnL"]

```

## Write function to plot transition probs based on input of trans matrix

```{r}
plot_transitions_input <- function(pars_list) {

  ns <- nrow(pars_list[[1]])  # number of states

  trans_prob <- data.frame(
    from = rep(1:ns, each=80*ns),
    to = rep(1:ns, each=80),
    PnL = seq(-20,60,length=80),
    prob = 0
  )

  y <- matrix(0.0,ncol=ns, nrow=80)

  for(from in 1:ns) {
    pars <- pars_list[[from]]
    for(to in 1:ns) {
      x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"PnL"]
      y[,to] <- exp(pars[to,1] + pars[to,2]*x)
    }
    y <- y/rowSums(y)
    for(to in 1:ns) {
      trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
    }
  }

  trans_prob <- trans_prob %>% mutate(from=recode(from, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust"),
                        to=recode(to, 
                         `1`="low-trust",
                         `2`="medium-trust",
                         `3`="high-trust")) %>% 
    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))

  ggplot(trans_prob,aes(x=PnL,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1))+
      theme(legend.position = "bottom")
}

```

```{r}
# pars has the format [[from]]["to","x"]
pars[[1]]["to_1","invPnL"] <- 0
pars[[1]]["to_1","intercept"] <-  10

pars[[2]]["to_1","intercept"] <- 3
#pars
plot_transitions_input(pars)
```