### 2024-02-13 screenCode.Rmd
more screening

---
title: "screeningAnalysis"
author: "Ismail Guennouni"
date: "2023-11-20"
output: pdf_document
---

```{r setup2, include=FALSE}
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(depmixS4)
library(lubridate)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data}
#rsq_dat <- read.csv("data/RSQ_Feb06_24.csv") 
rsq_dat <- read.csv("data/RSQ_Feb13_24.csv")

# Step 1: Remove rows 1 and 2
df_screening <- rsq_dat[-c(1, 2), ]

df_screening <- df_screening %>%
  mutate(
    MH = if_else(as.Date(ymd_hms(StartDate)) > as.Date("2023-11-28"), 1, 0),
    MH.f = factor(MH, levels = c(0, 1), labels = c("Gen pop", "MH Condition"))
  )

# Rename columns Q4 to Q12 to Q1 to Q9
old_names <- paste0("Q", 4:12)
new_names <- paste0("Q", 1:9)
names(df_screening)[names(df_screening) %in% old_names] <- new_names

# Rename the "Duration..in.seconds." column to "duration"
names(df_screening)[names(df_screening) == "Duration..in.seconds."] <- "duration"

# Step 2: Filter rows
df_screening <- df_screening[df_screening$Q1 == "I accept" & !grepl('test', df_screening$Q2) & df_screening$DistributionChannel ==  'anonymous', ]
cat("valid submissions: ",  nrow(df_screening), "\n")

# Calculate the number of valid responses (non-NA in all columns)
#valid_responses_count <- sum(complete.cases(df_screening))

###########################
# Step 3: Identifying responses where both Q_att_1 and Q_att_2 are less than or equal to 6
unsuccessful_responses <- df_screening$Q_att_1 < 6 | df_screening$Q_att_2 < 6

# Calculating the number of unsuccessful responses
unsuccessful_count <- sum(unsuccessful_responses)

# Print the number of unsuccessful responses
cat("Number of responses failing attention checks:", unsuccessful_count, "\n")

###########################

# # Convert duration to numeric (if it's not already)
# df_screening$duration <- as.numeric(as.character(df_screening$duration))
# 
# # Calculate the median duration for the whole dataset
# median_duration_total <- median(df_screening$duration, na.rm = TRUE)
# 
# # Calculate the median duration for those who failed the attention checks
# median_duration_unsuccessful <- median(df_screening$duration[unsuccessful_responses], na.rm = TRUE)
# 
# # Print the median durations
# cat("Median duration for the whole dataset:", median_duration_total, "seconds\n")
# cat("Median duration for unsuccessful responses:", median_duration_unsuccessful, "seconds\n")

#########################

# Filtering out these unsuccessful responses
df_screening<- df_screening[!unsuccessful_responses, ]

# Print the number of valid responses
cat("Number of valid responses:", nrow(df_screening), "\n")

```
```{r}
# Step 4: Calculate rejection sensitivity score
# Identifying columns ending with _1 and _2, ignoring Q_att_1 and Q_att_2
one_columns <- grep('_1$', names(df_screening), value = TRUE)
two_columns <- grep('_2$', names(df_screening), value = TRUE)
one_columns <- setdiff(one_columns, c('Q_att_1', 'Q_att_2'))
two_columns <- setdiff(two_columns, c('Q_att_1', 'Q_att_2'))

# The scale is from 1 to 6 for b columns, we reverse it by subtracting from 7
df_screening[two_columns] <- lapply(df_screening[two_columns], function(x) 7 - as.numeric(x))

# Calculate the rejection sensitivity score for each situation
for (i in 1:length(one_columns)) {
    score_col <- paste0(substr(one_columns[i], 1, nchar(one_columns[i])-2), '_score')
    df_screening[[score_col]] <- as.numeric(df_screening[[one_columns[i]]]) * as.numeric(df_screening[[two_columns[i]]])
}

# Step 5: Create a new column for the mean of the rejection sensitivity scores
score_columns <- grep('_score$', names(df_screening), value = TRUE)
df_screening$RS_score <- rowMeans(df_screening[, score_columns], na.rm = TRUE)

# Display the first few rows of the updated dataframe
head(df_screening)

```

```{r}

# Plot histogram of mean rejection sensitivity with overlaid density plot
ggplot(df_screening, aes(x = RS_score)) +
    geom_histogram(aes(y = after_stat(density)), binwidth = 1, fill = 'skyblue', color = 'black', alpha = 0.6) +
    geom_density(color = 'red', size = 1.5, alpha = 0.7) +
    theme_minimal() +
    labs(title = 'Histogram and Density of Mean Rejection Sensitivity Scores',
         x = 'Mean Rejection Sensitivity Score',
         y = 'Density') +
    theme(plot.title = element_text(hjust = 0.5),
          axis.title.x = element_text(size = 12, face = "bold"),
          axis.title.y = element_text(size = 12, face = "bold")) + 
   facet_wrap(~ MH.f)

```

```{r}
# Calculate the mean of the RS_score 
mean_score <- round(mean(df_screening$RS_score, na.rm = TRUE),2)

# Calculate the standard deviation
std_dev <- round(sd(df_screening$RS_score, na.rm = TRUE),2)

# Count the number of respondents with scores above 15 (inclusive)
count_above_15 <- sum(df_screening$RS_score >= 15, na.rm = TRUE)

# Count the number of respondents more than 1 standard deviation above the mean
count_above_1sd <- sum(df_screening$RS_score > (mean_score + std_dev), na.rm = TRUE)

# Count the number of respondents more than 2 standard deviations above the mean
count_above_2sd <- sum(df_screening$RS_score > (mean_score + 2 * std_dev), na.rm = TRUE)

# Print the results
cat("Mean of the mean rejection sensitivity scores:", mean_score, "\n")
cat("SD:", std_dev, "\n")
cat("Number of respondents with scores >= 15:", count_above_15, "\n")
cat("Number of respondents > 1 standard deviation above the mean:", count_above_1sd, "\n")
cat("Number of respondents > 2 standard deviations above the mean:", count_above_2sd, "\n")

```
```{r}
# Count the number of respondents with scores above 15 (inclusive) in both samples
count_above_15_normal <- round(
  (df_screening %>% dplyr::filter(MH == 0, RS_score >= 15) %>% nrow()) / 
  (df_screening %>% dplyr::filter(MH == 0) %>% nrow()) * 100, 1)

count_above_15_MH <- round(
  (df_screening %>% dplyr::filter(MH == 1, RS_score >= 15) %>% nrow()) / 
  (df_screening %>% dplyr::filter(MH == 1) %>% nrow()) * 100, 1)

cat("Proportion of HEALTHY respondents with scores >= 15 on RSQ :", count_above_15_normal, "%\n")
cat("Proportion of MH respondents with scores >= 15 on RSQ :", count_above_15_MH, "%\n")

```

```{r}
# Return Prolific IDs for those above a certain score:

high_RS <- df_screening %>% dplyr::filter(RS_score >= 15) %>% dplyr::select(PROLIFIC_PID)

high_RS

write.csv(high_RS, "data/high_RS.csv", sep = ",", row.names = FALSE,quote = FALSE)

```
```{r}
# Old high RS participants identified in screening in November/December 2023
high_RS_old <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) != as.Date("2024-02-06") ) %>% dplyr::select(PROLIFIC_PID)

# Newly identified 06 feb 2024
high_RS_new <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) == as.Date("2024-02-06") ) %>% dplyr::select(PROLIFIC_PID)

# String them together and add to prolific
new_ids_string <- paste(high_RS_new, collapse = ",")
ids_string
new_ids_string

```

```{r}
# Load csv of IDs of already collected data from high RSQ in early December
highRS_collected <- read.csv("data/highRSQ_collected.csv") 
```

```{r}
# Find IDs in high_RS that are not in highRS_collected
missing_ids <- !high_RS$PROLIFIC_PID %in% highRS_collected$Participant.id

# Subset high_RS to get the dataframe of people you haven't collected data for
not_collected <- high_RS[missing_ids, ]

nrow(not_collected)

ids_string <- paste(not_collected, collapse = ",")
ids_string
```

# Newly identified high RS on the 13th feb 2024.

```{r}

high_RS_13feb <- df_screening %>% dplyr::filter(RS_score >= 15, as.Date(ymd_hms(StartDate)) == as.Date("2024-02-13") ) %>% dplyr::select(PROLIFIC_PID)

nrow(high_RS_13feb)

# String them together and add to prolific
feb13_ids_string <- paste(high_RS_13feb$PROLIFIC_PID, collapse = ",")

feb13_ids_string
```

# Select those with low rejection sensitivity defined as below mean of 10 on the RSQ (...by the 6th of feb)
```{r}

# Return Prolific IDs for those above a certain score:

low_RS <- df_screening %>% 
          dplyr::filter(RS_score < 10) %>% 
          dplyr::select(PROLIFIC_PID)

# Extract the PROLIFIC_PID values as a character vector
low_RS_vector <- low_RS$PROLIFIC_PID

# Create the comma-separated string
low_RS_string <- paste(low_RS_vector, collapse = ",")

# Count the number of elements in the extracted PROLIFIC_PID vector
num_ids <- length(low_RS_vector)

cat("Number of extracted IDs: ", num_ids)

```