### 2024-03-27 efa.Rmd
v2 article

---
title: "EFA"
author: "Ismail Guennouni"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include =F)
library(tidyverse)
library(dplyr)
```

<!-- Load dataset 1 (RSQ) and select questionnaires -->
```{r rsq}
load("data/df_screening.RData")

# Identifying columns ending with _1 and _2, ignoring Q_att_1 and Q_att_2
one_columns <- grep('_1$', names(df_screening), value = TRUE)
two_columns <- grep('_2$', names(df_screening), value = TRUE)
one_columns <- setdiff(one_columns, c('Q_att_1', 'Q_att_2'))
two_columns <- setdiff(two_columns, c('Q_att_1', 'Q_att_2'))

df_rsq <- df_screening %>% dplyr::select(id, all_of(one_columns), all_of(two_columns))

```

<!-- Load dataset 2 (LPFS) and select questionnaires -->
```{r LPFS}

load("data/d_finished.RData")
df_lpfs <- d_finished %>% dplyr::select(id,contains("LPFS.")) %>% unique()

# Right join the data frames on the "id" column
df_efa <- merge(df_rsq, df_lpfs, by = "id", all.y = TRUE) %>% unique() %>% dplyr::select(-id)

```

```{r, include=T}
# Convert empty strings to NA and ensure numeric columns are indeed numeric.
# Replace the missing cells with column median
df_clean <- df_efa %>%
  mutate(across(everything(), ~as.numeric(as.character(.)))) %>%
  mutate(across(everything(), ~ifelse(is.na(.), median(., na.rm = TRUE), .)))

```

<!-- Duble checking non numeric and missing value cells -->
```{r}

str(df_clean) # all columns are numeric now 
colSums(is.na(df_clean)) # alls sums are 0 so no more missing values

```

```{r nFactors, include =T}

library(nFactors)
library(psych)

ev <- eigen(cor(df_clean)) # get eigenvalues

ev$values

scree(df_clean, pc=FALSE)  # Scree plot of the eigen values. Use pc=FALSE for factor analysis

```

```{r, include =T}
# another way to plot the eigenvalue
fa.parallel(df_clean, fa="fa")
```
The eigenvalue method (“Kaiser’s rule”) is telling us that 8 factors may be best. The scree plot is putting us somewhere between three and five factors. Parallel analysis suggests similar. THink 3 is adequate for such a small dataset but we can also use other methods such as CNG.

```{r, include =T}
Nfacs <- 3  # Think 3 factors is adequate for this small dataset. You can change this as needed.

# start with assumption of correlation btw factos. so using promax.
fit <- factanal(df_clean, Nfacs, rotation="promax")

print(fit, digits=2, cutoff=0.3, sort=TRUE)
```
```{r, include =T }
load <- fit$loadings[,1:2]
plot(load,type="n") # set up plot
text(load,labels=names(df_clean),cex=.7)
```
```{r, include =T}
loads <- fit$loadings

fa.diagram(loads)
```

```{r, include=T}
dim(fit$loadings) 
round(fit$loadings[ 1:30,], 2)
```

<!-- # Using the Polycor Package  -->

<!-- ```{r} -->
<!-- library(polycor) -->
<!-- library(nFactors) -->
<!-- hetcor_matrix <- hetcor(df_clean, polyserial = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ev_poly <- eigen(hetcor_matrix$correlations) # get eigenvalues -->

<!-- ev_poly$values -->

<!-- # Applying the CNG test -->
<!-- cng_results <- efa.CNG(hetcor_matrix$correlations, n.obs = nrow(df)) -->

<!-- # The CNG test results include the suggested number of factors to retain -->
<!-- print(cng_results) -->

<!-- ``` -->