### 2024-06-05 article.Rmd
version 3

---
title: "Can Exposure To Forgiving AI Foster Cooperative Play?"
author:
  # "Ismail Guennouni, Georgia Koppe, Christoph Korn"

keywords:
- Social Trust
- Repair of Cooperation
- Cognitive Manipulation 
- Economic Games 
- Hidden Markov Models

abstract: |

bibliography: "bib/exposure.bib"
biblio-style: spphys

output:
  bookdown::pdf_document2:
    toc: false
header-includes:
  - \usepackage{lscape}
---

```{r setupCoax, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, include=FALSE) 
knitr::opts_chunk$set(out.width = "\\textwidth")
```

```{r load-packages05}
library(papaja)
library(kableExtra)
require(knitr)
#require(citr)
require(bookdown)

# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(grid)
library(gridExtra)
library(forcats)
library(ggsignif)
library(magick)

```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(42)
options(tinytex.verbose = TRUE)
```

```{r emmOptions}
emm_options(pbkrtest.limit = 6210)
emm_options(lmerTest.limit = 6210)

```

# Introduction

<!-- Trust-based cooperation is an important element of human social interactions, facilitating seamless interpersonal and intergroup relations -->

<!-- <!-- ### Lack of trust and social dysfunction  -->

<!-- Research into the determinants of psychopathology has also linked trust-based constructs to the emergence of mental health disorders [@fonagy_mentalizing_2017]. Many studies have drawn attention to difficulties in social cognition as a crucial concern within personality disorders (See review in @herpertz_social-cognitive_2014). Individuals with these disorders often find it challenging to initiate and sustain social interactions, leading to behaviors that undermine their relationships. Specifically, individuals facing such social cognitive challenges tend to act uncooperatively, which has been identified as a key marker for the severity and presence of  a cluster of Personality Disorder (PD) symptoms [@mulder_relationship_1999]. -->

<!-- <!-- ### Could lack of trust be due to sampling bias for uncaring early caregivers ? -->

<!-- One perspective of how these difficulties in the social domain emerge is through experience with early caregivers. Attachment theory [@bowlby_attachment_1978] posits that the quality of early caregiver relationships shapes an individual's future ability to form secure attachments and trust others. Research strongly suggests a connection between insecure attachment styles (characterized by anxiety or avoidance) and lower levels of trust in relationships. Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998]. Similarly, @fonagy_role_2014 posit that when a young learner grows up surrounded by unreliable communicators, they may adapt by becoming mistrustful of social knowledge. This mistrust serves as a protective mechanism, particularly within environments where caregivers lack benign intention.  -->

<!-- <!-- In their model, drawn on Bowlby’s (1973) attachment theory, Downey and colleagues (1996, 1998) proposed that RS may evolve through early experiences of repeated rejection from primary caregivers, which lead individuals to form the expectation that they are likely to be rejected by others. This expectation is associated with hypervigilance for rejection-relevant cues, which in turn leads to negative emotional responses and maladaptive behavioral reactions. Even if RS is more strongly activated in threatening situations, it is assumed to be a stable disposition, a sort of implicit theory about the self (i.e., rejectable) and others (i.e., rejecters). As such, it can lead to long-term negative effects on individual functioning (e.g., self-regulation, mental health, per- sonality, social cognition) and undermine relationship quality (Gao, Assink, Cipriani, & Lin, 2017).  -->

<!-- <!-- ### Exposure as a way to correct the sampling bias.  -->

<!-- If this adaptive mistrust is the source of social dysfunction, we can ask whether exposing those who exhibit it to cooperative and forgiving interaction partners might correct this bias.  Research in the fields of behavioral economics and psychology has explored how positive social interactions influence trust and cooperation. The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions [@joyce_trust_1995]. In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors and trust levels, even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005]. @fowler_cooperative_2010 studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior. Similarly, research on social learning theory [@bandura_social_1977] has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior. These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust. -->

<!-- <!-- ### Describe experiment -->

<!-- In this study, using a randomised control trial design, we assess the effectiveness of a manipulation aimed at repairing the potential breakdown of cooperation in the RTG from a low investment sent by a computerized investor. The manipulation focuses on repeatedly exposing participants to agents that have been designed to have a limited propensity to retaliate to a non-cooperative action by the participants. In order to select a sample that might exhibit difficulties in building relationships based on Trust dues to adverse early experiences, we focused on stratifying participants based on Rejection Sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection. It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997]. Extensive research demonstrates strong associations between rejection sensitivity and various mental health conditions, including depression, anxiety, personality disorders, and self-harm (for a review, see @gao_associations_2017). However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas. By startifying participants based on RS, we can directly examine whether the manipulation is particularly effective  for those most vulnerable to trust-related issues. Additionally, comparing outcomes for those with high and low RS will help us determine if the exposure-based manipulation has differential impacts depending on an individual's underlying sensitivity to rejection. The computerized investor in the RTG was programmed to play according to a hidden Markov model estimated from real players' data in prior research. A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios. Such a model, informed by empirical data and integrating a responsive trust mechanism, represents a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control. -->

Trust is fundamental to human social interactions, facilitating seamless relations at both interpersonal and intergroup levels.  The study of psychopathology has linked deficits in trust-based constructs to the development of mental health disorders [@fonagy_mentalizing_2017].  Individuals with personality disorders often struggle to form and maintain social connections, a difficulty reflected in uncooperative behaviors – a marker for the severity of PD symptoms [@herpertz_social-cognitive_2014; @mulder_relationship_1999].

One explanation for such social challenges lies in early caregiver experiences. Attachment theory [@bowlby_attachment_1978] suggests that the quality of these relationships shapes our capacity for secure attachments and trust. Individuals with higher levels of insecure attachment may recall negative trust-related experiences more easily, report fewer positive trust experiences, and use less constructive coping strategies when trust is broken [@mikulincer_attachment_1998]. These insecure attachment patterns are often associated with heightened rejection sensitivity (RS), a tendency to anxiously expect, readily perceive, and intensely react to rejection [@downey_early_1997]. RS has been linked to the development of various mental health conditions, including depression, anxiety, personality disorders, and self-harm [@gao_associations_2017]. Given the potential influence of RS on trust-related dynamics, it becomes particularly relevant to our understanding of social dysfunction. Could an ingrained sensitivity to rejection contribute to a mistrust bias? Similarly, learners exposed to unreliable communicators could develop mistrust of social knowledge as a protective strategy [@fonagy_role_2014].

If this adaptive mistrust is the source of social dysfunction, we can ask whether exposing those who exhibit it to cooperative and forgiving interaction partners might correct this bias.  Research in the fields of behavioral economics and psychology has explored how positive social interactions influence trust and cooperation. The use of the repeated trust game (RTG), a well-established experimental approach, has allowed for the analysis of the development of trust through ongoing interactions [@joyce_trust_1995]. In this paradigm, cycles of mutual trust, where each party's trust is reciprocated with trustworthiness, have the effect of enhancing cooperative behaviors and trust levels, even among individuals who are initially inclined to be distrustful [@king-casas_getting_2005]. @fowler_cooperative_2010 studied behavior in social networks interacting in a public goods game and found that cooperative behavior tends to cluster, suggesting that exposure to cooperative peers can lead to more cooperative behavior. Similarly, research on social learning theory [@bandura_social_1977] has long demonstrated that individuals learn and model the behavior of those around them, indicating that if someone is consistently exposed to cooperative and positive individuals, they're likely to emulate this behavior. These insights highlight that engaging with compassionate and forgiving others can be an effective method for mitigating deeply ingrained mistrust.

<!-- In this study, we use a randomized control trial to test whether exposing participants to forgiving and more cooperative co-players can produce more trustworthy behavior and repair potential breakdowns in RTG cooperation. We stratify the sample of participants based on rejection sensitivity (RS). In order to simulate realistic social interaction while maintaining a high degree of experimental control, the computerized investor in the RTG was designed to play using a generative model of how humans play. We fit Hidden markov Models (HMMs) to real players' data from prior studies to build a responsive yet consistent co-player. A key aspect of this agent is that its actions depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios. An advantage of having a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation. Such generative  model of behavior, informed by empirical data and integrating a responsive trust mechanism, represent a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control. In this study, we leverage this flexibility to expose participants in between trust games to artificial agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust.  In the control condition, the participants play the human like agents throughout.  We would expect those in the manipulation condition to behave more cooperatively and have a lower propensity to retaliate to the co-player's defection. We would also expect those exhibiting higher rejection sensitivity to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity. -->

In this study, we use a randomized control online experiment to test whether exposing participants with varying levels of RS to forgiving and more cooperative co-players results in more trustworthy behavior and a repair of potential breakdowns in RTG cooperation. In order to simulate realistic social interaction while maintaining a high degree of experimental control, we take a novel paradigmatic approach: We use generative models of how humans play the RTG to design an agent that plays the role of the investor, based on Hidden Markov Models (HMMs) fitted to real players data. A key aspect of these agents is that their actions depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios. An advantage of having such a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation. In this study, participants were given the role of the trustee. After playing a 15 rounds RTG with a human-like HMM investor, they were randomly assigned to either a Control or Manipulation condition. In the Manipulation condition, participants were exposed over three RTGs to HMM investors designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust. In the Control condition, participants, played three RTGs against the same human-like HMM. After this exposure phase, all participants played another 15 round RTG with a human-like HMM investor.  We hypothesise that those in the manipulation condition would behave more cooperatively and have a lower propensity to retaliate to the co-player's defection after the exposure phase. We also expect those exhibiting higher RS to be more responsive to the pre-programmed change in the agent's cooperativeness and retaliation propensity.

# Methods

```{r loadCSV}

library(dplyr)
d_finished <- readRDS("../scriptMongo/data/d_finished.csv")

# Step 1: Identify the expo_opponent for each playerId where gameNumber == "11" (exposure game) & isLastRound == TRUE (to select a particular)
expo_opponent_df <- d_finished %>%
  filter(gameNumber == "11", isLastRound == TRUE) %>%
  dplyr::select(id, expo_opponent = game_opponent) %>%
  distinct() # Ensure uniqueness in case of duplicates

# Step 2: Join this information back to the original dataset
d_finished <- d_finished %>%
  left_join(expo_opponent_df, by = "id") %>%
  mutate(condition.f = factor(expo_opponent, levels= c("AI_HMM","AI_HMM_nice"),
                              labels=c("Control","Manipulation"))) %>%
  dplyr::select(-expo_opponent) # remove the expo_opponent column if it's no longer needed

###### get whether high or low RS and score from df_screening

# Load the "df_screening.RData" file
load("data/df_screening.RData")

RS_data <- df_screening %>%  mutate(high_RS = factor(ifelse(RS_score > 15, 1, 0), 
                                    levels = c(0, 1), 
                                    labels = c("low RS", "high RS")))  %>% 
                             dplyr::select(id, MH.f, RS_score, high_RS)

# Find IDs for those who participated in experiment and add RS_score and high_RS as new columsn to d_finished.
# Merging d_finished with relevant columns from RS_data based on matching ids
d_finished <- d_finished %>%
  inner_join(RS_data %>% dplyr::select(id, RS_score, high_RS), by = "id")

save(d_finished, file = "data/d_finished.RData")

# Summarize to count the number of participants for each Condition within each high_RS level
temp <- d_finished %>% dplyr::select(id, condition.f, high_RS) %>% 
                      unique() %>%
                      group_by(high_RS, condition.f) %>%
                      summarize(participant_count = n(), .groups = 'drop')

# View the summarized counts
print(temp)
```

## Participants

A total of 206 participants (56% female) were recruited on the Prolific Academic platform (prolific.co). The mean age of participants was 34.6 years, with an 11.9 years standard deviation. Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5. Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire (RSQ) to form two similarly sized groups: One with high RS (RSQ score > 15) and the other with low RS (RSQ score < 10).

## Design and Procedure

The experiment had a 2 (Condition: Manipulation or Control) by 2 (RS : high or low) by 2 (Game: Trust-Game Pre Manipulation, Trust-Game Post Manipulation) design, with repeated measures on the third factor (Figure \@ref(fig:HMMPanels).A). Participants within each pre-screened group were randomly assigned to one of the two levels of the first factor. The games were designed and implemented online using Empirica v1 [@almaatouq_empirica_2021].

## Tasks and Measures

```{r, include=FALSE}
Anti_social <-  c(0.1025436430408, 0.1221931236853, 0.1345215238995,
0.1368182252617, 0.1285592471751, 0.1116014322677, 0.0895041220882,
0.0663166721334, 0.0453950277118, 0.0287077419587, 0.0167723588011,
0.0090530028158, 0.0045143317507, 0.0020796744990, 0.0008851094702,
0.0003480141146, 0.0001264134789, 0.0000424213859, 0.0000131513231,
0.0000037665630, 0.0000009965755)

Neutral <- c(0.003393529, 0.006930874, 0.013053553, 0.022671228,
0.036310144, 0.053627606, 0.073039389, 0.091734929, 0.106248217,
0.113479701, 0.111769796, 0.101517390, 0.085028780, 0.065675083,
0.046778251, 0.030725245, 0.018610323, 0.010394861, 0.005354126,
0.002543094, 0.001113881)

Pro_social <- c(0.003162697, 0.001057162, 0.001410197, 0.001881016,
0.002508877, 0.003346113, 0.004462480, 0.005950950, 0.007935434,
0.010581067, 0.014107909, 0.018809194, 0.025075644, 0.033427845,
0.044559370, 0.059394206, 0.079163228, 0.105506029, 0.140606514,
0.187373417, 0.249680651)

investment <- seq(0:20) -1

response_probs <- as.data.frame(cbind(investment,Anti_social,Neutral,Pro_social)) %>% 
  rename("low-trust" = "Anti_social", "medium-trust"="Neutral", "high-trust" = "Pro_social") %>%
  pivot_longer(cols=c("low-trust","medium-trust","high-trust"),
                    names_to='Investor_state',
                    values_to='probability') %>% 
   mutate(across(Investor_state, factor, levels=c("low-trust","medium-trust","high-trust")))
```

```{r, include=F}

plotinvHMM <- ggplot(response_probs,                            
       aes(x = investment,
           y = probability,
           fill = Investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state', x = "Investment", y= "Probability") + 
  theme_bw() + 
  theme(legend.position = "bottom",  legend.text = element_text(size = 12),  legend.title = element_text(size = 14))

plotinvHMM
```

```{r, include=FALSE}

# Parameters for HMM human-like Transition Function
unhappy_pars <- rbind(c(0,0), c(-3.366027, 0.40910797 ), c(-3.572619,0.08137274)) 
neutral_pars <- rbind(c(0,0), c(3.3142637, 0.3763408), c(0.9169736, 0.4502838))
happy_pars   <- rbind(c(0,0), c(0.7134085, 0.02101626), c(2.2215478 ,0.16162964))

pars_inv <- list(unhappy_pars, neutral_pars, happy_pars)

# Parameters for HMM nice Transition Function
unhappy_pars_nice <- rbind(c(-10,0), c(-3.366027, 0.40910797 ), c(-3.572619,0.08137274)) 
neutral_pars_nice <- rbind(c(0,0), c(3.3142637, 0.3763408), c(0.9169736, 0.4502838))
happy_pars_nice   <- rbind(c(0,0), c(0.7134085, 0.02101626), c(2.2215478 ,0.16162964))

pars_inv_nice <- list(unhappy_pars_nice, neutral_pars_nice, happy_pars_nice)

plot_HMM_transitions <- function(ns, pars_mat) {

  trans_prob <- data.frame(
    from = rep(1:ns, each=100*ns),
    to = rep(1:ns, each=100),
    ret = seq(-20,60,length=100),
    probs = 0
  )

  y <- matrix(0.0,ncol=ns, nrow=100)

  for(from in 1:ns) {
  pars <- matrix(pars_mat[[from]], ncol=2)
  # print(pars)

    for(to in 1:ns) {
        x <- trans_prob[trans_prob$from == from & trans_prob$to == to,"ret"]
        y[,to] <- exp(pars[to,1] + pars[to,2]*x)
    }
    y <- y/rowSums(y)

    for(to in 1:ns) {
      trans_prob$probs[trans_prob$from == from & trans_prob$to == to] <- y[,to]
    }
  }

  df <- as.data.frame(trans_prob) %>% 
    mutate(from = recode(from, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust"),
           to = recode(to, "1" = "low-trust", "2" = "medium-trust", "3" = "high-trust") ) %>% 
    mutate(across(from, factor, levels=c("low-trust","medium-trust","high-trust"))) %>% 
    mutate(across(to, factor, levels=c("low-trust","medium-trust","high-trust")))

    # Create a separate data frame with the background colors
  bg_colors <- data.frame(
    from = factor(c("low-trust", "medium-trust", "high-trust"), levels=c("low-trust","medium-trust","high-trust"))
  )

  # plotting code...
  ggplot() +
    geom_rect(data = bg_colors, aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = from), alpha = 0.1) +
    geom_line(data = df, aes(x = ret, y = probs, colour = as.factor(to))) +
    facet_wrap(~from, labeller = labeller(from = function(x) paste("From", x, "state on trial t"))) +
    ylim(c(0,1)) +
    scale_fill_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
                    name = "From state") +  # Changed legend title for 'fill' here
    scale_color_manual(values = c("low-trust" = "red", "medium-trust" = "green", "high-trust" = "blue"),
                       labels = c("low-trust", "medium-trust", "high-trust"),
                       name = "State transitioned to") +
    labs(x = "Investor's net return on trial t", y = "Transition probability to \nState on trial t+1", color = 'State transitioned to') +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.key.size = unit(1, 'lines'),
          legend.spacing.x = unit(0.1, 'in'),
          legend.title = element_text(size = 14),
          legend.margin = margin(t = 0.2, b = 0, unit = 'cm'),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
          strip.text = element_text(size = 10),
          legend.box = "vertical" # Arrange legends vertically
    ) +
    guides(fill = guide_legend(order = 1, title.position = "left", title.hjust = 0.3),
           color = guide_legend(order = 2, title.position = "left", title.hjust = 0.3))
}

```

```{r, include=F}

plotInvTran <- plot_HMM_transitions(3, pars_inv) 
print(plotInvTran)

plotInvTranNice <- plot_HMM_transitions(3, pars_inv_nice) 
print(plotInvTranNice)

```

### Repeated Trust Game

Participants played a 15-round RTG [@joyce_trust_1995] in the trustee role against a computer-programmed investor. On each round the investor is endowed with 20 units and decides how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this tripled amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. Each player was represented with an icon with the participant always on the left of the screen and the co-player on the right. The participants were able to choose the icon that represents them at the start of the experiment. The icon representing the co-player changed at the start of each new game, to simulate a new interaction partner. Participants were not told they were facing computerised co-players.

\begin{landscape}
```{r HMMPanels, fig.cap= "\\small{A: A timeline of the experiment for both conditions. The RTG is played in dyads, with participants always assigned the role of the trustee and the HMM agent that of the investor. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee who needs to decide how much of the multiplied investment they want to send back to the investor. The difference between conditions is the type of agents participants are exposed to in the exposure phase. Panels B - D: We construct the artificial investor agent by fitting a three-state hidden Markov model to data of human investors engaged in the 10 round RTG against human trustees. From the fitted HMM, we get the distribution of investments by the human-like agent, conditional on its latent state as shown in Panel B. The fitted HMM also yields the transition probability of the agent to a state on trial t+1 as a function of the net return (difference between the investment sent and the amount received in return) on trial t as shown in Panel C. Each plot in Panel C represents a different starting latent state on trial t, and each line represents the probability of transitioning to a particular state in trial t+1. Panel D shows the transition probabilities of the forgiving HMM agent which are identical to those of the human-like agent, except in the low trust state. Unlike the human-like HMM, the forgiving HMM always transitions out of the low-trust state, and is more likely to end up in a high-trust state.}",include=T, out.width='1\\linewidth', out.height='1\\textheight', fig.align='center'}

library(magick)

# Read in the images
image1 <- image_read("figures/timeline.png")
image2 <- image_read("figures/HMMinvPolicy2.png")
image3 <- image_read("figures/HMMinvTrans2.png")
image4 <- image_read("figures/HMM_switch.png")

# Determine the target width for all images, for example, the width of the widest image
target_width <- max(image_info(image1)$width, image_info(image2)$width, 
                    image_info(image3)$width, image_info(image4)$width)

target_height <- max(image_info(image1)$height, image_info(image2)$height, 
                    image_info(image3)$height, image_info(image4)$height)

# Scale images to have the same width
image1 <- image_scale(image1, paste0(target_width, "x", target_height))
image2 <- image_scale(image2, paste0(target_width, "x", target_height))
image3 <- image_scale(image3, paste0(target_width, "x", target_height))
image4 <- image_scale(image4, paste0(target_width, "x", target_height))

# Annotate each image with a label
image1 <- image_annotate(image1, text = "A", location = "+0+10", size = 90, color = "black", gravity = "northwest")
image2 <- image_annotate(image2, text = "B", location = "+0+10", size = 90, color = "black", gravity = "northwest")
image3 <- image_annotate(image3, text = "C", location = "-0+10", size = 90, color = "black", gravity = "northwest")
image4 <- image_annotate(image4, text = "D", location = "+0+10", size = 90, color = "black", gravity = "northwest")

# Assuming a 300 DPI resolution for good print quality
dpi <- 300

# Spacing of 0.1 inches converted to pixels
spacing <- 0.05 * dpi

# Calculating the width and height for each image
width_per_image <- ((12 - 0.1) / 2) * dpi
height_per_image <- ((7.4 - 0.1) / 2) * dpi
# The geometry string for each image, including spacing (assuming no borders)
geometry_string <- paste0(width_per_image, 'x', height_per_image, '+', spacing, '+', spacing)

# Now create the montage with the specified geometry
combined_image <- image_montage(
  c(image1, image2, image3, image4), 
  tile = "2x2",
  geometry = geometry_string
)

# Display the combined image
plot(combined_image)
```
\end{landscape}

The strategy of the computerised investor was modelled on the behavior of human investors in the RTG using an existing dataset consisting of 388  ten round games with the same co-player. Full detail on the datasets used in the Supplementary Information. Using this data, we estimated a hidden Markov model (HMM) on investors' behavior with three latent states. The HMM consists of three outputs. First, a distribution over the possible investments from 0 to 20 for each latent state of the investor (Figure \@ref(fig:HMMPanels).B). These distributions can be interpreted as reflecting "low-trust", "medium-trust", or "high-trust" dispositions. The HMM also specifies how  the investor can move between these latent states.  The probability of these transitions was modelled as a function of their net return (i.e return - investment) in the previous round (see Figure \@ref(fig:HMMPanels).C). Finally, the HMM also provides as an output the initial distribution of latent states for the investor. 
In order to instigate a potential breakdown of trust, thereby allowing us to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 (pre-manipulation) or round 13 (post-manipulation). On all other rounds, the investor's actions were determined by randomly drawing an investment from the state-conditional distribution, with the state over rounds determined by randomly drawing the next state from the state-transition distribution as determined from the net return on the previous round (disregarding the net return immediately after the pre-programmed low investment rounds). The initial state for the HMM investor in each instance of the game was the "mid-trust" state.

## Manipulation

The HMM resulting from fitting the model to human investor play in the RTG (Human-like HMM) isused as a basis to design a forgiving and ultimately more cooperative agent. To achieve that, we changed the transition function of the investor HMM to make it impossible to remain in a low trust state once the agent transitions there. This is done by selecting  parameters of the transition function to make the probability of remaining in the "low-trust" state, when the agent is in that state, effectively nil. The resulting transition function is shown in Figure \ref{fig:HMMPanels}.D. The policies conditional on the latent states and the transition function in the other latent states remain unchanged.

## Procedure

At the start of the experiment, participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Participants were randomly assigned to either a Control or Manipulation condition. The timeline of the experiment is shown in Figure \@ref(fig:HMMPanels).A.  Phase one ("pre") consisted of a 15 round RTG in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor participants decided how much of the tripled investment to return to the investor, before continuing to the next round. After completing 15 rounds of the RTG, participants rated how cooperative, forgiving they perceived the investor to be, and whether they would like to play with them again (all on a scale from 1 to 10 with 10 being the most positive rating).

Phase 2 ("exposure") consisted of three 7-round RTGs. Participants in the Manipulation condition faced the forgiving HMM investor and rated the agent on the same attributes as in the pre-manipulation phase.  Those in the Control condition faced the same human-like HMM agent as in the "pre" phase and rated each co-player on the same attributes. To keep the experience similar to the "pre" phase, the agent in the control condition was also designed to send a very low investment in round 5 of each of the three games. In the post-manipulation phase ("post"), participants in both conditions were told they would face a new player, and faced the same human-like HMM as in "pre" phase. Participants then completed the Levels of Personality Functioning Scale (LPFS) questionnaire (see the supplement for details). Finally, participants were asked whether they thought the other players were human or computer agents, then debriefed and thanked for their participation.

## Statistical analysis

To explore whether participants behaved differently in the RTG after the manipulation compared to the control group, we model the percentage return (percentage of tripled investment returned to investor) using a linear mixed-effects model with Phase (RTG game pre vs. post-manipulation), Condition (manipulation vs. control), Investment, and RS (High vs Low RS group) as well as their interactions as fixed effects, and player-wise random intercepts and slopes for Phase.

The model was estimated using the `afex` package [@singmann_afex_2022] in R. More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek_balancing_2017]. A similar process was used to establish the random effects structures of  linear mixed-effects models used to analyse the HMM agent investments as well as the participants' ratings of the co-players. For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex". We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.

```{r ratingsProcess}

# d_finished now has the 'Condition' column with the value equal to expo_opponent for all rounds and rows per playerId

d_finished <- d_finished  %>% mutate( gameNum.f = factor(gameNumber,labels = c("pre", "expo1", "expo2", "expo3","post"),
                                      levels=c("1", "11", "12", "13", "2")),
                                      phase.f = factor(ifelse(gameNumber %in% c("1", "2"), 
                                                       ifelse(gameNumber == "1", "pre", "post"), NA), levels = c("pre", "post"))) %>% 
                              rename(# Before Exposure
                            rating_cooperative_1 = `ratePLayer_trustGame_1.How cooperative do you think the other player is?`,
                            rating_forgiving_1 = `ratePLayer_trustGame_1.How forgiving do you think the other player is?`,
                            rating_playAgain_1 = `ratePLayer_trustGame_1.If given the choice, would you like to play again with the other player?`,

                            # During Exposure
                            rating_cooperative_11 = `ratePLayer_trustGame_11.How cooperative do you think the other player is?`,
                            rating_forgiving_11 = `ratePLayer_trustGame_11.How forgiving do you think the other player is?`,
                            rating_playAgain_11 = `ratePLayer_trustGame_11.If given the choice, would you like to play again with the other player?`,
                            rating_cooperative_12 = `ratePLayer_trustGame_12.How cooperative do you think the other player is?`,
                            rating_forgiving_12 = `ratePLayer_trustGame_12.How forgiving do you think the other player is?`,
                            rating_playAgain_12 = `ratePLayer_trustGame_12.If given the choice, would you like to play again with the other player?`,
                            rating_cooperative_13 = `ratePLayer_trustGame_13.How cooperative do you think the other player is?`,
                            rating_forgiving_13 = `ratePLayer_trustGame_13.How forgiving do you think the other player is?`,
                            rating_playAgain_13 = `ratePLayer_trustGame_13.If given the choice, would you like to play again with the other player?`,

                            # After Exposure
                            rating_cooperative_2 = `ratePLayer_trustGame_2.How cooperative do you think the other player is?`,
                            rating_forgiving_2 = `ratePLayer_trustGame_2.How forgiving do you think the other player is?`,
                            rating_playAgain_2 = `ratePLayer_trustGame_2.If given the choice, would you like to play again with the other player?`,

                            )
```

```{r datRatings}

#######################    Pre-processing player ratings

# Pivot longer ratings on each attribute 
df_coop <- d_finished %>% dplyr::select(playerId,condition.f,contains("cooperative"),high_RS) %>%  
  pivot_longer(cols=contains("cooperative"), names_to = c("game"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()

df_forgiv <- d_finished %>% dplyr::select(playerId,condition.f,contains("forgiving"),high_RS) %>%
  pivot_longer(cols=contains("forgiving"), names_to = c("game"), names_pattern ="rating_forgiving_(.*)", values_to = "rating_forgiving") %>% distinct()

df_playAgain <- d_finished %>% dplyr::select(playerId,condition.f,contains("again"),high_RS) %>%  
  pivot_longer(cols=contains("again"), names_to = c("game"), names_pattern ="rating_playAgain_(.*)", values_to = "rating_playAgain") %>% distinct()

#merge all data frames together
datRatings <- list(df_coop, df_forgiv,df_playAgain) %>% 
              reduce(full_join, by=c('playerId','game','condition.f','high_RS')) %>%
              mutate(gameNum.f = factor(game,labels = c("pre", "expo1", "expo2", "expo3","post"),
                                      levels=c("1", "11", "12", "13", "2")))

# Group means for each rating 
mu <- datRatings %>% group_by(gameNum.f, condition.f) %>% summarise(mean_coop = mean(rating_coop),
                                                  mean_forgiv = mean(rating_forgiving),
                                                  mean_playAgain = mean(rating_playAgain)
                                                  )
mu

# ggplot(mu,aes(x=gameNum.f)) + 
#   geom_bar(aes(y=mean_forgiv, stat = "identity", position = "dodge", color="blue")) + 
#   geom_bar(aes(y=mean_coop, stat = "identity", position = "dodge", color="darkgreen")) +
#   geom_bar(aes(y=mean_playAgain, stat = "identity", position = "dodge", color="red")) +
#   ggtitle("Player ratings after each game") +
#   xlab("Game Number") +
#   theme_minimal() + 
#   theme(legend.position = "bottom")
#   
```

```{r LPFS}
library(dplyr)

# Step 1: Convert LPFS columns to numeric
d_finished <- d_finished %>%
  mutate(across(starts_with("LPFS"), as.numeric))

# Step 2: Calculate means
# Identify LPFS columns
lpfs_cols = grep("^LPFS", names(d_finished), value = TRUE)
lpfs_1_to_6_cols = grep("^LPFS\\.[1-6]$", names(d_finished), value = TRUE)
lpfs_7_to_12_cols = grep("^LPFS\\.(7|8|9|10|11|12)$", names(d_finished), value = TRUE)

# Calculate means using vectorized operations
d_finished <- d_finished %>%
  mutate(
    LPFS_score = rowMeans(dplyr::select(., all_of(lpfs_cols)), na.rm = TRUE),
    LPFS_self_score = rowMeans(dplyr::select(., all_of(lpfs_1_to_6_cols)), na.rm = TRUE),
    LPFS_other_score = rowMeans(dplyr::select(., all_of(lpfs_7_to_12_cols)), na.rm = TRUE)
  )

```

```{r dfCorr}

library(dplyr)
library(GGally)

df_corr <- d_finished %>% dplyr::select(id, RS_score, LPFS_score, LPFS_other_score,payoffTrust1,payoffTrust2) %>%
  unique()

# df_corr %>% summarise(
#   Pearson_correlation = cor(LPFS_score, RS_score, method = "pearson", use = "complete.obs"),
#   Spearman_correlation = cor(LPFS_score, RS_score, method = "spearman", use = "complete.obs")
# )
# 
# # View the correlations
# print(correlations)
# 
# # Now plot the scatter plot for the entire dataset (not by individual id)
# ggplot(df_corr, aes(x = LPFS_score, y = RS_score)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE) +
#   labs(x = "LPFS Score", y = "RS Score", title = "Scatterplot of LPFS Score vs. RS Score") +
#   theme_minimal()

# Select only the numerical columns for the correlation matrix (excluding the 'id' column)
numerical_data <- df_corr %>% dplyr::select(-id)

# Calculate the correlation matrix
correlation_matrix <- cor(numerical_data, use = "complete.obs")

# Print the correlation matrix
print(correlation_matrix)

# Now create a scatterplot matrix
# Install the package if you haven't already
if (!require("GGally")) install.packages("GGally")

ggpairs(numerical_data)

```

```{r corrPayoffs}

temp2 <- d_finished %>% dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f, RS_score, LPFS_score) %>%
  mutate(perf = payoffTrust1 + payoffTrust2) %>% 
  unique()

# #                            Bonus Calculation
# avg_perf <- mean(temp2$perf)
# lowRS_bonus <- temp2 %>% filter(high_RS== "low RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
# write.csv(lowRS_bonus,"data/bonusLow.csv")
# 
# highRS_bonus <- temp2 %>% filter(high_RS== "high RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
# write.csv(highRS_bonus,"data/bonusHigh.csv")

ggplot(data = temp2, aes(x = payoffTrust1, y = payoffTrust2)) +
  geom_point() +  # Color points by id
  labs(x = "Payoff Trustee 1", y = "Payoff Trustee 2", color = "ID") +
  theme_minimal() +
  theme(legend.position = "right")

```

```{r}
# Sanity checks

# prolific_highRS <- read.csv(file="/Users/mbp17/Downloads/prolific_export_highRS.csv")
# 
# # Which Ids in highRS file but not in Prolific? 
# 
# # Find IDs in highRS_bonus that are not in prolific_highRS
# ids_in_highRS_not_in_prolific <- setdiff(highRS_bonus$id, prolific_highRS$Participant.id)
# 
# # Find IDs in prolific_highRS that are not in highRS_bonus
# ids_in_prolific_not_in_highRS <- setdiff(prolific_highRS$Participant.id, highRS_bonus$id)
# 
# # ids_in_highRS_not_in_prolific now contains IDs that are in highRS_bonus but not in prolific_highRS
# # ids_in_prolific_not_in_highRS now contains IDs that are in prolific_highRS but not in highRS_bonus
# 
# # print these variables to see the IDs
# print(ids_in_highRS_not_in_prolific)
# print(ids_in_prolific_not_in_highRS)
```

```{r}
library(ggplot2)
library(dplyr)

# First, reshape data into a longer format.
long_data_temp <- d_finished %>%
  dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
  dplyr::select(id, payoffTrust1, payoffTrust2) %>%
  gather(key = "game", value = "payoff", -id) %>%
  mutate(game = factor(game, levels = c("payoffTrust1", "payoffTrust2"))) %>% 
  unique()

# Now, create the dumbbell plot
ggplot(long_data_temp, aes(x = game, y = payoff, group = id)) +
  geom_line(aes(group = id), color = "gray") + # Connect payoffs with a line
  geom_point(aes(color = id), size = 3, alpha = 0.6) + # Payoff points
  theme_minimal() +
  theme(legend.position = "none") + # Hide legend if too many playerids
  labs(x = "Game", y = "Payoff", title = "Payoff Comparison between Game 1 and Game 2")

```

```{r}

df_long_ratings <- datRatings %>%
  pivot_longer(
    cols = starts_with("rating_"), 
    names_to = "rating_type", 
    values_to = "rating_value"
  ) %>%
  dplyr::select(playerId, condition.f, gameNum.f, rating_type, rating_value, high_RS)

# First, calculate the mean and SEM for each group
summary_df <- df_long_ratings %>%
  group_by(gameNum.f, condition.f, rating_type, high_RS) %>%
  summarise(
    mean = mean(rating_value, na.rm = TRUE),
    sem = sd(rating_value, na.rm = TRUE)/sqrt(n())
  ) %>%
  ungroup()

```

<!-- ## Linear mixed effects Model for returns  -->

```{r avgRetDf}
# Filter only trust rounds and create % returns and investments

avg_ret_df <- d_finished %>%
  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS, LPFS_score, RS_score) %>% 
  filter(roundType=="trust",!is.na(roundNum)) %>% 
  mutate(roundNum = as.numeric(as.character(roundNum))) %>%
  mutate(inv_scaled = as.vector(scale(investment))) %>% 
  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)),ret_pct_na = ifelse(investment==0,NA,returns/(3*investment)))

# Number of people left for analysis. (there are 51 rounds in each game)
#num_participants <- nrow(avg_ret_df)/51
num_participants <- length(unique(avg_ret_df$playerId))

num_participants

```

```{r}
# average investment for first 10 rounds pre manipulation
avg_pct <- avg_ret_df %>% filter(roundNum < 11 , gameNum.f == "pre") %>% summarise(avg_inv = mean(inv_pct), avg_ret = mean(ret_pct_na, na.rm=TRUE))
avg_pct
```

```{r}
# Plot Invs and Rets in absolute values for pre and post
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=roundNum)) + 
  stat_summary(aes(y=returns, color="Returns"), fun = "mean", geom = "line", group = 1) +
  stat_summary(aes(y=investment, color="Investment"), fun = "mean", geom = "line", group = 1) +
  facet_wrap(~gameNum.f) +
  scale_color_manual(values = c("Returns" = "red", "Investment" = "blue")) +
  labs(color = "") +
  xlab("Round Number")+
  theme_bw() +
  theme(legend.position = "bottom",  # Move legend to bottom
        legend.title.align = 0.5,  # Center-align the legend title
        legend.key = element_blank(),
        plot.title = element_text(hjust = 0.5),  # Center-align the plot title
        axis.text = element_text(color = "black"),  # Customize axis text
        axis.title = element_text(color = "black")) + # Customize axis title
  facet_wrap(~condition.f*gameNum.f)

# # Plot Invs and Rets in percentages for pre and post
# ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")),aes(x=roundNum)) + 
#   stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=ret_pct_0)) + 
#   stat_summary(fun = "mean", geom = "line",group=1, color = "blue", alpha =0.6,position = position_dodge2(width=0.8),aes(y=inv_pct)) + 
#   geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
#   facet_wrap(~gameNum.f)

ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=roundNum)) + 
  stat_summary(aes(y=ret_pct_0, color="Returns"), fun = "mean", geom = "line", group = 1) +
  stat_summary(aes(y=inv_pct, color="Investment"), fun = "mean", geom = "line", group = 1) +
  geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
  facet_wrap(~gameNum.f) +
  scale_color_manual(values = c("Returns" = "red", "Investment" = "blue")) +
  labs(color = "") +
  xlab("Round Number")+
  ylab("Investment / Returns (%)")+
  theme_bw() +
  theme(legend.position = "bottom",  # Move legend to bottom
        legend.title.align = 0.5,  # Center-align the legend title
        plot.title = element_text(hjust = 0.5),  # Center-align the plot title
        axis.text = element_text(color = "black"),  # Customize axis text
        axis.title = element_text(color = "black"))+ # Customize axis title
  facet_wrap(~condition.f*gameNum.f)

```

```{r}
# Plot Invs and Rets in absolute values for exposure
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3")),aes(x=roundNum)) + 
  stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=returns)) + 
  #stat_summary(fun.data = "mean_ci", geom = "errorbar",alpha =0.4, width = 0.2, aes(y=ret_pct) ) + 
  stat_summary(fun = "mean", geom = "line",group=1, color = "blue",aes(y=investment)) + 
  facet_wrap(~condition.f*gameNum.f)

# Plot Invs and Rets in percentages for exposure
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("expo1","expo2","expo3")),aes(x=roundNum)) + 
  stat_summary(fun = "mean", geom = "line",group =1, color = "red", aes(y=ret_pct_0)) + 
  #stat_summary(fun.data = "mean_ci", geom = "errorbar",alpha =0.4, width = 0.2, aes(y=ret_pct) ) + 
  stat_summary(fun = "mean", geom = "line",group=1, color = "blue", alpha =0.6,position = position_dodge2(width=0.8),aes(y=inv_pct)) + 
  geom_hline(yintercept=0.33, linetype="dashed", color = "grey") +
  facet_wrap(~condition.f*gameNum.f)

```

<!-- Linear mixed effects model for all returns  -->

# Behavioral Results

## Analysis of participant returns

```{r modAllReturns, include=FALSE, cache=TRUE}

prepost_df <- avg_ret_df %>% filter(gameNum.f %in% c("pre","post"))

mod_returns_pct_RS <- mixed( ret_pct_na ~ phase.f*condition.f*inv_scaled*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")

summary(mod_returns_pct_RS)

```

```{r}
saveRDS(summary(mod_returns_pct_RS), "data/mod_returns_pct_RS.RDS")
```

```{r}
contrasts(prepost_df$phase.f)
contrasts(prepost_df$condition.f)
contrasts(prepost_df$high_RS)

```
```{r}
#RS_model <- lm( RS_score ~ ret_pct_na*inv_scaled + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
```

```{r prepostByCond, cache=TRUE}
prepost_bycond <- pairs(emmeans::emmeans(mod_returns_pct_RS, c("phase.f"), by = "condition.f", pbkrtest.limit = 6300))
prepost_bycond
```

```{r modemmeans, cache=TRUE, include=F}
library(emmeans)

# Compute EMMs for the interaction of interest
emms <- emmeans(mod_returns_pct_RS, ~ phase.f * condition.f | high_RS, pbkrtest.limit = 6300)

contrasts <- contrast(emms, "pairwise", by = c("condition.f", "high_RS"))
summary(contrasts)

results_df <- as.data.frame(summary(contrasts))
```

<!-- EFFECT SIZES -->
```{r}
# Load the necessary packages

library(sjstats)
effectsize::eta_squared(mod_returns_pct_RS)
```

On average, investments and returns, as shown in Figure \ref{fig:gamesPlot}, fell within the documented range of 40-60% of the endowment for investments and 35-50% of the total yield for returns, as reported in previous studies [@charness_investment_2008; @fiedler_social_2011].

```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received (y-axis) by Condition, Phase, and game round (x-axis). The blue line shows the returns in the Pre phase and the green line those in the Post phase. The left Panel shows returns in the Control condition and the right one those in the Manipulation condition. The dotted lines identify the rounds where the pre-programmed one-off low investment occurs. We note lower average returns post vs pre in the manipulation condition, whilst returns in the control condition are similar between the two phases.",fig.align="center", fig.width=6, fig.height = 4}

# Create a data frame for the vertical linesr
vline_data <- data.frame(
  xintercept = c(12, 13),
  Defection_round = factor(c("Pre-manipulation", "Post-manipulation"), levels = c("Pre-manipulation", "Post-manipulation"))
)

# Plot
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
               alpha = 0.3, linetype = 0) +
  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Trust Game Phase") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "vertical", # This will stack the legends vertically
        legend.key = element_blank(),
        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_color_manual(values = c("darkblue", "darkgreen"),
                     labels = c("Pre-manipulation", "Post-manipulation")) +
  scale_fill_manual(values = c("darkblue", "darkgreen"),
                    guide = "none") + # Hide the fill legend
  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
         linetype = guide_legend(title = "Defection round")) +
  facet_wrap(~condition.f)

```

Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs. Post RTG game), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with higher percentage returns in the Pre RTG compared to the Post. Importantly, we also find an interaction between Condition and Phase (RTG pre- vs. post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`. As shown in Figure \ref{fig:boxPlots}, post-hoc tests confirm that, contrary to our expectations, there was a decrease in the percentage returned only in the manipulation condition, pre - post,
`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition.  We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.

There was also a significant main effect of Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$inv_scaled)`,
such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, reflected  that returns were more affected by investments in the control condition. We also find a three way interaction between Phase,  Investment and RS
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation). Finally, we find a four-way interaction between Condition, Phase, Investment and RS `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`. For completeness, these effects are discussed in more detail the supplement.

<!-- # USING LMER TO LEVERAGE CAPABILITIES WITH EMTRENDS -->

```{r}
library(lmerTest)
library(emmeans)

# Refitting the model directly
direct_model <- lmer(ret_pct_na ~ phase.f * condition.f * inv_scaled * high_RS + (1 + phase.f | playerId), data = prepost_df, REML = TRUE)

emm_res <- emmeans(direct_model, specs = ~ phase.f * condition.f * high_RS * inv_scaled)
trends <- emtrends(direct_model, specs = ~ phase.f * condition.f * high_RS, var = "inv_scaled")

# Print results
summary(emm_res)
summary(trends)

```

```{r fourwayInteractionPlot, include=F, fig.cap ="Visualising the four way interaction between Phase, RS group, Condition and Investment. The plot shows the effect of the investment on participant returns for each condition, phase and RS group. The interesting effect is in the Manipulation Condition: " }

library(ggplot2)

# 
# ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
#   # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
#   geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
#   labs(title = "Effect of inv_scaled on Outcome by Condition",
#        x = "inv_scaled", y = "Outcome") +
#   theme_minimal() +
#   facet_wrap(~high_RS)

# 'inv_scaled' is the continuous predictor, and 'outcome' is your dependent variable.
ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
  # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
  geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
  labs(x = "Investment Z-values", y = "Return proportion", color = "Experiment Phase") +
  theme_minimal() +
  facet_wrap(~condition.f*high_RS)+
  theme(legend.position = "bottom")

```

<!-- To interpret this interaction we plot the effect of the Investment on returns for each level of Condition, Phase and RS. Figure \ref{fig:fourwayInteractionPlot} shows that while the effect of investment on returns is similar across RS groups and across phases in the Control condition, the Manipulation condition shows a higher effect of investment in the post phase compared to pre in the low RS group, but a lower effect of investment compared to before for the high RS group. This suggests that high RS participants based their return less on the investment after being exposed to the forgiving AI compared to the low RS group. -->

```{r invMod, cache=TRUE}

mod_invs_RS <- mixed( investment ~ phase.f*condition.f*high_RS + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")

summary(mod_invs_RS)

```

```{r violPlotsDef, cache=TRUE}
library(afex)
library(ggplot2)
library(ggsignif)

# Generate the plot with points and error bars
ret_all <- afex_plot(mod_returns_pct_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
                     mapping = c("linetype", "shape", "fill"),
                     factor_levels = list(condition.f = c("Control", "Manipulation")),
                     legend_title = "Phase")

# Extract the data used in afex_plot
plot_data <- ret_all$data

# Generate the plot
ret_all <- ggplot(plot_data, aes(x = condition.f, y = y, color = phase.f, group = phase.f)) +
  geom_point(position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = y - SE, ymax = y + SE), width = 0.2, position = position_dodge(width = 0.8)) +
  theme_bw() +
  labs(y = "% Returns", x = "Condition", color = "Trust Game Phase") +
  theme(plot.title = element_text(size = 16), legend.position = "bottom") +
  geom_signif(
              annotations = "***", 
              xmin = 1.8, xmax = 2.2,
              y_position = 0.47, 
              tip_length = 0.03, 
              textsize = 3, 
              color = "black")

######################################################################################################

# Generate the plot with points and error bars
inv_all <- afex_plot(mod_invs_RS, x = "condition.f", trace = "phase.f", dodge = 0.8, error = "within",
                     mapping = c("linetype", "shape", "fill"),
                     factor_levels = list(condition.f = c("Control", "Manipulation")),
                     legend_title = "Phase")

# Extract the data used in afex_plot
plot_data <- inv_all$data

# Generate the plot
inv_all <- ggplot(plot_data, aes(x = condition.f, y = y, color = phase.f, group = phase.f)) +
  geom_point(position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = y - SE, ymax = y + SE), width = 0.2, position = position_dodge(width = 0.8)) +
  theme_bw() +
  labs(y = "HMM Investment", x = "Condition", color = "Trust Game Phase") +
  theme(plot.title = element_text(size = 16), legend.position = "bottom")+
  ylim(9, 11)

```

```{r boxPlots, include=T,fig.cap="Marginal means and distributions of either investments or percentage returns across participants by Phase and Condition. The top panel shows that participants in the Intervention condition returned lower proportions of the multiplied investment received in the second game compared to the first game over all ronds, whilst those in the Control condition sent back similar returns. The bottom panel shows no difference on aggregate of how the HMM invested across Phases and Conditions.",fig.align="center", fig.width=6, fig.height = 8}

library(patchwork)

ret_all/inv_all +  plot_layout(guides = 'collect') & theme(legend.position = "bottom")

```

<!-- Focusing on post defection only:  -->
### Post Defection Trials

```{r postDefMod, cache=TRUE}
post_def_data <- avg_ret_df %>% filter(roundType=="trust",!is.na(gameNum.f),(roundNum >= 12 & gameNum.f =="pre")| (roundNum >= 13 & gameNum.f =="post"))

mod_retPostDef_RS <- mixed( ret_pct_na ~ phase.f*condition.f*inv_scaled + (1 + phase.f| playerId), post_def_data , REML= TRUE, method="KR")

summary(mod_retPostDef_RS)
```

```{r PairsPostDef}
phasePairs <- pairs(emmeans::emmeans(mod_retPostDef_RS, c("phase.f"), pbkrtest.limit = 1449))
phasePairs
```

Did participants learn to be more forgiving and cooperative after witnessing the pre-programmed defection by the HMM investor?. To explore this question, we restrict the analysis to the trials following the pre-programmed defection by the HMM agent in both the "pre" (trials 12 to 15) and the "post" phases (trials 13 to 15).  We fit the same mixed effects model as for all the trials with the exception of the RS variable. This is because RS did not show main or interaction effects in the main model, and also due to the necessity of running a simpler model to accommodate the low number of trials. We find a significant main effect of Phase `r papaja::apa_print(mod_retPostDef_RS)$full_result$phase_f` with returns lower in the second game post defection trials compared to the first. We also find a main effect of Investment 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_retPostDef_RS)$full_result$inv_scaled)` 
where participants continued to return higher proportions when receiving higher investments. 
Finally, we still find an Investment by Condition interaction 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)` showing a lower effects of investment on the manipulation condition compared to the control condition in post-defection trials. However, the absence of a Condition by Phase effect indicates there was no difference between conditions on participants' reaction to a one-off defection by the co-player.

<!-- Checking whether Investor behavior has changed dramatically between pre and post: -->

```{r}
library(ggplot2)

# Plotting
ggplot(prepost_df %>% filter(condition.f=="Manipulation", high_RS == "high RS", roundNum < 12), aes(x = roundNum, y = investment, group = playerId, color = as.factor(gameNum.f))) +
  geom_line(alpha = 0.5) +  # Set alpha for transparency to handle overlap
  facet_wrap(~ playerId, scales = 'free_y') + # Faceting by playerID, each gets its own y-axis scale
  theme_minimal() +
  labs(color = 'Game Number') +
  theme(
    strip.text = element_blank(), # Remove facet labels since there will be too many
    axis.text.x = element_text(angle = 90, hjust = 1) # Rotate x-axis labels for readability
  )

```

### HMM investor
<!-- Linear mixed effects model for investments -->
To test whether the HMM agent's behavior differed between Phases, Conditions and RS groups, we estimate a linear mixed-effects model of investments sent by the computerised HMM agent with Condition, Phase and RS and their interaction as fixed effects, and a similar random effects structure to the returns model. As seen in Figure \@ref(fig:boxPlots), we find no main or interaction effects, indicating HMM behavior was on aggregate similar across Phase, Conditions and RS groups.

When asked during debrief whether they thought the investors they faced were Human or not, $41$% of participants thought they were either facing a human or were not sure of the nature of the co-player. When asked to justify their choice, many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial co-player's behavior.

### Questionnaire scores and performance

<!-- We found no correlation between either the RS score and task performance, as defined by the total points accumulated in both the "pre" and "post phase of the RTG. The LPFS score was also not correlated with task performance. These two scores exhibited a positive correlation  (Rho = 0.52, p < 0.001) -->

```{r}
cor.test(temp2$LPFS_score, temp2$perf, method = "spearman")
cor.test(temp2$RS_score, temp2$perf, method = "spearman")

cor_result <- cor.test(temp2$RS_score, temp2$LPFS_score, method = "spearman")
# Format the result using papaja
apa_result <- papaja::apa_print(cor_result)

print(apa_result)
# Extract the correlation coefficient and p-value
spearman_rho <- apa_result$estimate

```

Whilst we found a significant correlation between participant's Levels of Personality Functioning Score (LPFS) and the Rejection Sensitivity score (RSQ), Spearman's `r spearman_rho`, $p < 0.001$, there was no correlation between these questionnaire scores and participant's return or overall task performance.

<!---------------------- Modelling of co-player ratings -------------------------------------->

```{r afexModCoop, cache=TRUE}

mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")

summary(mod_rating_coop)

```

```{r emmCoop, cache=TRUE}
# Contrast 1: Pre vs. Post
contrast1 <- c(-1, 0, 0, 0, 1)

# Contrast 2: Pre vs. (Expo1, Expo2, Expo3) 
contrast2 <- c(-3, 1, 1, 1, 0)

# Contrast 3: Post vs. (Expo1, Expo2, Expo3) 
contrast3 <- c(0, -1, -1, -1, 3)

# Combine contrasts into a matrix
contrast_matrix <- rbind(contrast1, contrast2, contrast3)
colnames(contrast_matrix) <- levels(datRatings$gameNum.f)

# Apply contrasts to the model using 'emmeans'
emm_coop <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f)

# Custom comparisons: use 'contrast' directly
comp1_coop <- contrast(emm_coop, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
comp2_coop <- contrast(emm_coop, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
comp3_coop <- contrast(emm_coop, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")

comp1_coop

comp2_coop

comp3_coop

```

```{r}
# Compute the estimated marginal means (EMMs) for the interaction
emm_coop_interaction <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f * high_RS)

comp1_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Post" = contrast_matrix[1,]))
comp2_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
# comp3_coop_interaction <- contrast(emm_coop_interaction, list("Post vs Expo" = contrast_matrix[3,]))

# Output the results
print("Pre vs Post")
comp1_coop_interaction
print("Pre vs Expo")
test(comp2_coop_interaction, adjust="bonferroni")
test(comp2_coop_interaction)

# test_comp3_coop_interaction
```

```{r afexModForg}

mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")

#summary(mod_rating_forg)

```

```{r emmForg}

# Apply contrasts to the model using 'emmeans'
emm_forg <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f)

# Custom comparisons: use 'contrast' directly
comp1_forg <- contrast(emm_forg, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
comp2_forg <- contrast(emm_forg, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
comp3_forg <- contrast(emm_forg, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")

comp1_forg

comp2_forg

comp3_forg

```

```{r}
# Compute the estimated marginal means (EMMs) for the interaction
emm_forg_interaction <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f * high_RS)

comp1_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Post" = contrast_matrix[1,]))
comp2_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
# comp3_forg_interaction <- contrast(emm_forg_interaction, list("Post vs Expo" = contrast_matrix[3,]))

# Output the results
print("Pre vs Post")
comp1_forg_interaction
print("Pre vs Expo")
comp2_forg_interaction

# test_comp3_forg_interaction
```

```{r afexModAgain}

mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")

#summary(mod_rating_again)

```

```{r emmAgain}

# Apply contrasts to the model using 'emmeans'
emm_again <- emmeans(mod_rating_again, ~ gameNum.f | condition.f)

# Custom comparisons: use 'contrast' directly
comp1_again <- contrast(emm_again, list("Pre vs Post" = contrast_matrix[1,]), by = "condition.f")
comp2_again <- contrast(emm_again, list("Pre vs Expo" = contrast_matrix[2,]), by = "condition.f")
comp3_again <- contrast(emm_again, list("Post vs Expo" = contrast_matrix[3,]), by = "condition.f")

comp1_again

comp2_again

comp3_again

```

```{r}
# Compute the estimated marginal means (EMMs) for the interaction
emm_again_interaction <- emmeans(mod_rating_again, ~ gameNum.f | condition.f * high_RS)

comp1_again_interaction <- contrast(emm_again_interaction, list("Pre vs Post" = contrast_matrix[1,]))
comp2_again_interaction <- contrast(emm_again_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
# comp3_again_interaction <- contrast(emm_again_interaction, list("Post vs Expo" = contrast_matrix[3,]))

# Output the results
print("Pre vs Post")
comp1_again_interaction
print("Pre vs Expo")
comp2_again_interaction

# test_comp3_again_interaction
```

## Player ratings

Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition and RS group. We will focus on two contrasts to analyse the ratings by Condition and RS group. The first is between the rating in the first phase ("pre") when participants phase the human-like HMM, and the average rating during the exposure phase (average of "expo1", "expo2" and "expo3") where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition). The second contrast is between the "pre" and "post" phases of the experiment where in both conditions participants face the same human-like HMM.

### Comparing pre and exposure ratings

For those with high RS, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`. There was no difference in ratings on forgiveness and whether they would like to face the co-players again. 
Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.

For those with low RS, there was no difference in any of the ratings between the "pre" and "exposure" phases for the Manipulation condition. For the Control condition, participants indicated less willingness to face the exposure co-player compared to the "pre" player, `r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo` but also did not differ on cooperation and forgiveness ratings. In summary, those with low RS had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving. In contrast, we see that exposing participants with high RS to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.

```{r plotRatings,include=T, echo=FALSE, fig.pos='H', fig.cap="Averages and standard errors of the participants ratings of the opponent (y-axis) by each game and condition for each phase (x-axis). Pre and post are the 15 round RTGs before and after the exposure phase respectively. The games titled expo 1 to 3 are the three 7 round games during the exposure phase. The blue line represents participants' perception of co-player cooperation, the red line indicates perceived co-player forgiveness, and the green line shows the participants' willingness to play again with the same co-player. Cooperation, forgiveness, and willingness to play again ratings remained relatively stable for the low RS group except post-manipulation, where they were lower. In the high RS group, the ratings more accurately reflected the agent's actual cooperativeness and forgiveness in the manipulation condition, but decreased over time in the control group,. Ratings were also markedly lower in the post-manipulation game for this group.",fig.align="center", fig.width=6, fig.height = 6}

ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f), color = rating_type)) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1, position = position_dodge(width = 0.25)) +
  geom_line(linetype = "dashed", position = position_dodge(width = 0.25)) +
  geom_point(position = position_dodge(width = 0.25), size = 3) +
  facet_wrap(~high_RS*condition.f, scales = "free_x") +
  labs(x = "Phase", y = "Mean Rating", color = "Rating Type") +
  scale_color_manual(values = c("rating_coop" = "lightblue", "rating_forgiving" = "#FF6961", "rating_playAgain" = "#77DD77"),
                     labels = c("Cooperative", "Forgiving", "Play again")) +
  theme_bw() +
  theme(legend.position = "bottom")

```

```{r}
# temporarily paste code from Figure 2 to compare styles: 
# Plot
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
               alpha = 0.3, linetype = 0) +
  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Trust Game Phase") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "vertical", # This will stack the legends vertically
        legend.key = element_blank(),
        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_color_manual(values = c("darkblue", "darkgreen"),
                     labels = c("Pre-manipulation", "Post-manipulation")) +
  scale_fill_manual(values = c("darkblue", "darkgreen"),
                    guide = "none") + # Hide the fill legend
  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
         linetype = guide_legend(title = "Defection round")) +
  facet_wrap(~condition.f)

```

```{r}
library(ggplot2)

ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f))) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem, color = condition.f), 
                width = .1, position = position_dodge(width = 0.25)) +
  geom_line(aes(color = condition.f), linetype = "dashed", position = position_dodge(width = 0.25)) +
  geom_point(aes(color = condition.f), position = position_dodge(width = 0.25), size = 3) +
  facet_grid(high_RS ~ rating_type, scales = "free_x") +
  scale_color_manual(values = c("Control" = "blue", "Manipulation" = "red")) +
  labs(x = "Phase", y = "Mean Rating", color = "Condition") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

### Comparing pre and post ratings
<!-- Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
<!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`. Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.  -->

Participants high on RS in the Manipulation condition rated the co-players in the "post" phase similarly on cooperation, lower on 
forgiveness `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
and lower on willingness to face them again `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`. 
Those in the Control condition rated the investors post the exposure phase lower on all three attributes 
(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`, 
Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).

For participants low on RS, those in the Manipulation condition rated the co-players in the "post" phase lower on all three attributes 
(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`). 
Those in the Control condition did not differ in their ratings of the "pre" and "post" phase players.

In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the control condition. In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the control condition even though the player continues to use the same strategy. After exposure to the forgiving HMM, both groups rate the "post" co-player worse than the "pre" even though they are the same.

```{r}
d_finished %>% 
  dplyr::select(id,Turing.choice) %>% 
  unique() %>% 
  group_by(Turing.choice) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)
```

<!-- Extracting Investor latent state from data -->

```{r extractInvStates}

library(tidyverse)

# Convert dataset from wide to long format, focusing on columns that start with "State_"
d_long <- d_finished %>%
  gather(key = "state_col", value = "investorState", starts_with("State_")) %>%
  mutate(
    round_extracted = as.numeric(str_extract(state_col, "(?<=State_r)\\d+")),
    game_extracted = as.numeric(str_extract(state_col, "(?<=game_)\\d+"))
  ) %>%
  # Filter to keep rows where extracted round/game numbers match the actual round/game numbers
  filter(round_extracted == roundNum, game_extracted == gameNumber) %>%
  # Select and rename the desired columns
  dplyr::select(id, condition.f, gameNum.f, gameNumber, roundNum, investorState)

# View the resulting dataset
print(d_long)

```

```{r InvStatePlot}
library(ggplot2)
library(dplyr)

# Filter for only games 1 and 2
filtered_data <- d_long %>%
  filter(gameNum.f %in% c("pre", "post"))

# Define the colors for the levels of investorState

state_colors <- c("happy" = "Blue", "neutral" = "#77DD77", "unhappy" = "#FF6961")

# Create the histogram with stacked bars and specified colors
ggplot(filtered_data, aes(x = as.factor(roundNum), fill = as.factor(investorState))) +
  geom_bar(aes(group = investorState), position = position_stack(reverse = FALSE), stat = "count") +
  scale_fill_manual(values = state_colors) +
  facet_wrap(~condition.f*gameNum.f, scales = "free_x", labeller = label_both) +
  labs(x = "Round", fill = "Investor State", title = "Histogram of Latent States per Game and Round") +
  theme_bw() +
  theme(legend.position = "bottom")

```

```{r}
counts <- filtered_data %>%
  group_by(condition.f, gameNum.f, investorState,id) %>%
  summarise(count = n()) %>%
    summarise(count = sum(count))

# Calculate the average count per state per condition
total_rounds <- filtered_data  %>%
  group_by(condition.f, gameNum.f) %>%
  summarise(total_rounds = n())

state_summary  <- counts %>%
  left_join(total_rounds, by = c("condition.f", "gameNum.f")) %>% 
  mutate(state_percentage = count/total_rounds)

ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = as.factor(investorState))) +
  geom_bar(stat = "identity") +
  facet_wrap(~condition.f) +
  labs(x = "Condition", y = "Proportion", fill = "Investor State") +
  theme_bw() +
  theme(legend.position = "bottom")

state_colors <- c("happy" = "Blue", "neutral" = "#77DD77", "unhappy" = "#FF6961")

# Create the histogram with stacked bars and specified colors
ggplot(state_summary , aes(x = gameNum.f, y = state_percentage, fill = as.factor(investorState))) +
  geom_bar(aes(group = investorState), position = position_stack(reverse = FALSE), stat = "identity") +
  scale_fill_manual(values = state_colors) +
  facet_wrap(~condition.f*gameNum.f, scales = "free_x", labeller = label_both) +
  labs(x = "Round", fill = "Investor State", title = "Histogram of Latent States per Game and Round") +
  theme_bw() +
  theme(legend.position = "bottom")
```

# Discussion

We used a randomized control online experiment where participants played a RTG with AI agents designed to simulate human-like trust-building scenarios. Participants were then exposed to either forgiving or human-like AI agents before playing another RTG. Contrary to our hypothesis, exposure to forgiving agents did not increase trust or cooperation. Instead, participants reduced their returns overall whilst the returns of those in the control group did not change between the pre and post phase of the experiment. Neither did participants show more forgiving behavior in the manipulation condition after the one-off defection by the agent. Why did participants reduce their returns even though they were repeatedly exposed to a more cooperative and more forgiving AI? A look at how the participants rated their co-players might shed some light on what might be driving this reduction in returns for those in the manipulation condition.

Those exposed to the forgiving agent rated their opponent in the post-exposure phase lower on all attributes even though they faced the same dynamic human-like HMM as pre-exposure. One possible explanation for this drop in rating is that participants exhibited a negative contrast effect. This occurs when the evaluation of a person, object, or situation is influenced by comparisons with recently encountered contrasting objects or people. If we've recently interacted with someone exceptionally nice, our perception of a normal level of niceness might be skewed, making normal behavior seem less favourable or even negative by comparison [@kobre_negative_1972]. As the most recently faced opponents were more forgiving and cooperative, this negative contrast effect may have trumped any learning transfer from being repeatedly exposed to cooperative and forgiving AI [@zentall_within-trial_2005]. If this contrast effect is indeed replicable, then an avenue for future research would be to use it to our benefit by making the participants play agents with low cooperation perception.

<!-- Participants also perceived the forgiving AI during the exposure phase as more cooperative, but not more forgiving. The latter might be the result of the participants not having many opportunities to test the co-player's forgiveness propensity since most decided to continue cooperating. -->

It is worth highlighting that RS was not correlated to overall performance in the task.  Neither did it moderate the change in returns between Conditions or have an effect on participant returns. However, in examining the player ratings by RS group more closely, we observe that individuals with high RS demonstrate a heightened attunement to changes in the behavior of their AI co-players. Specifically, when exposed to a more forgiving AI agent, these participants accurately increased their ratings of the agent, indicating a sensitive and appropriate response to the behavioral manipulation. This adjustment reflects a nuanced perception of social cues and a capacity to modify judgments based on the behavior of interaction partners, potentially indicating a perceived alignment or support from the co-player that mitigates their heightened sensitivity to potential social rebuffs. High RS participants' lower ratings of essentially similar co-players as they continued to face them may indicate that the occasional pre-programmed defections of the AI agents failed to mitigate concerns over rejection and possibly exacerbated perceptions of uncooperativeness and unforgiveness, leading to a stronger preference against future interactions. 
Individuals with low RS were less affected by AI behavioral variations, showing a less volatile baseline of social perception. They still exhibit behavior consistent with a contrast effect in the manipulation condition. This dichotomy highlights the potential of tailored interactions, mediated by advanced AI agents, to address and modulate specific social and psychological predispositions in human participants.

<!-- Interestingly, for participants with low RS, the manipulation did not significantly alter their perceptions of the AI's cooperativeness, forgiveness, or their willingness to interact in the future. This group's ratings remained largely unchanged between the pre- and post-exposure phases, regardless of the AI's behavior. This stability suggests that individuals with lower sensitivity to rejection may base their ratings more on consistent behavioral patterns or other criteria not directly influenced by the forgiving nature of the AI agents. -->

This study leveraged a novel approach to examining the dynamics of trust and cooperation in social interactions through the utilization of generative models of human behavior to design artificial agents that can interact in economic games. The use of these agents in the RTG led to similar investment and returns to those recorded in human dyadic interactions. Participants were often uncertain whether they interacted with human or artificial investors, highlighting the agents’ realism. This validates the use of these artificial agents to probe the effectiveness of manipulations whilst keeping a high degree of experimental control. Importantly, our approach utilizing HMMs to simulate interactive partners in social dilemmas offers a promising avenue for future research, especially in understanding and potentially mitigating trust deficits in individuals with high RS. The nuanced behaviors elicited through the interaction with HMM agents highlight the complexity of trust dynamics and the potential for computational models to offer novel insights into human social behaviors.

<!-- In conclusion, this study demonstrates the value of integrating HMM-based artificial agents in the study of trust and cooperation. By showcasing the potential of these agents to both simulate realistic human interactions and influence real human behavior, our research paves the way for future investigations into the mechanisms of trust repair and the broader implications of artificial agents in social psychology. -->

<!-- ## Other notable points: -->

<!-- The study found that RSQ (RS Questionnaire) and LPFS (Levels of Personality Functioning Scale) scores were ineffective at distinguishing between high and low performers in the game. This gap highlights a need for further research, despite theoretical connections suggested in the literature review. -->

\pagebreak

# References