### 2024-06-26 exploratory.Rmd
trstee HMM fit & joint state analytics

---
title: "exploratory"
author: "Ismail Guennouni"
date: "`r Sys.Date()`"
output: pdf_document
---

Here we fit an HMM to all trustee data in the exposure experiment, both pre/post and control/manipulation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(depmixS4)
library(flextable)
library(grid)
library(gridExtra)
library(forcats)
library(ggsignif)
library(magick)
library(doParallel)

```

## Exploratory analyses to model individual games

```{r}
load("data/d_finished.RData")

data_HMM <- d_finished %>%
  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS, LPFS_score, RS_score) %>% 
  filter(roundType=="trust",!is.na(roundNum),gameNum.f %in% c("pre","post")) %>% 
  mutate(roundNum = as.numeric(as.character(roundNum))) %>%
  mutate(inv_scaled = as.vector(scale(investment))) %>% 
  mutate(next_investment = lead(investment, default=0)) %>%
  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)),ret_pct_na = ifelse(investment==0,NA,returns/(3*investment)))
```

```{r}

# Define a response class representing Gaussian distributions with discrete support
setClass("discgaus", contains="response", slots=c(breaks="numeric"))

# Define a generic function for creating instances of the discgaus class
setGeneric("discgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("discgaus"))

# Define a method for creating instances of the discgaus class
setMethod("discgaus", 
          signature(y="ANY"), 
          function(y, pstart=NULL, fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), ...) {
            # Convert y to a matrix
            y <- matrix(y, length(y))
            # Create x matrix
            x <- matrix(1)
            # Initialize parameters list
            parameters <- list()
            # Number of parameters
            npar <- 2
            # Check if fixed parameters are provided, else initialize as FALSE
            if(is.null(fixed)) fixed <- as.logical(rep(0, npar))
            # Check if initial parameters are provided, else initialize default values
            if(!is.null(pstart)) {
              if(length(pstart) != npar) stop("length of 'pstart' must be ", npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            } else {
              parameters <- list(mu = 10, sigma = 3)
            }
            # Create a new object of class discgaus
            mod <- new("discgaus", parameters=parameters, fixed=fixed, x=x, y=y, npar=npar, breaks=breaks)
            mod
          }
)

# Define method for displaying parameters of discgaus class
setMethod("show", "discgaus",
          function(object) {
            cat("Gaussian with discrete support\n")
            cat("Parameters: \n")
            cat("mu: ", object@parameters$mu, "\n")
            cat("sigma: ", object@parameters$sigma, "\n")
          }
)

# Define method for computing density of discgaus class
setMethod("dens", "discgaus",
          function(object, log=FALSE) {
            p <- pnorm(object@breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - 
                 pnorm(object@breaks[-length(object@breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
            if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) 
            else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
          }
)

# Define method for setting parameters of discgaus class
setMethod("setpars", "discgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values) != npar) stop("length of 'values' must be", npar)
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

# Define method for getting parameters of discgaus class
setMethod("getpars", "discgaus",
          function(object, which="pars", ...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

# Define method for fitting discgaus class
setMethod("fit", "discgaus",
          function(object, w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object, c(pars[1], exp(pars[2])))
            object
          }
)

# Define a response class representing truncated Gaussian distributions with discrete support
setClass("truncdiscgaus", contains="discgaus", slots=c(min="numeric", max="numeric"))

# Define method for computing density of truncdiscgaus class
setMethod("dens", "truncdiscgaus",
          function(object, log=FALSE) {
            breaks <- c(object@min, object@breaks[object@breaks > object@min & object@breaks < object@max], object@max)
            prec <- pnorm(object@max, mean = object@parameters$mu, sd = object@parameters$sigma) - 
                    pnorm(object@min, mean = object@parameters$mu, sd = object@parameters$sigma)
            if(is.na(prec) || prec < 1e-12) {
              p <- rep(1/(length(breaks)-1), length(object@y))
              if(log) return(log(p)) else return(p)
            } else {
              p <- pnorm(breaks[-1], mean = object@parameters$mu, sd = object@parameters$sigma) - 
                   pnorm(breaks[-length(breaks)], mean = object@parameters$mu, sd = object@parameters$sigma)
              p <- p/sum(p)
              if(log) return(log(p[as.numeric(cut(object@y, breaks=object@breaks))])) 
              else return(p[as.numeric(cut(object@y, breaks=object@breaks))])
            }
          }
)

# Define a generic function for creating instances of the truncdiscgaus class
setGeneric("truncdiscgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("truncdiscgaus"))

# Define a method for creating instances of the truncdiscgaus class
setMethod("truncdiscgaus", 
          signature(y="ANY"), 
          function(y, pstart=NULL, fixed=NULL, breaks = c(-Inf, seq(0, 19) + .5, Inf), min=-0.5, max=20.5, ...) {
            y <- matrix(y, length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0, npar))
            if(!is.null(pstart)) {
              if(length(pstart) != npar) stop("length of 'pstart' must be ", npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            }
            mod <- new("truncdiscgaus", parameters=parameters, fixed=fixed, x=x, y=y, npar=npar, breaks=breaks, min=min, max=max)
            mod
          }
)

# Define a response class representing Gaussian distributions with variable discrete support
setClass("vtdgaus", contains="response", slots=c(yield="numeric"))

# Define a generic function for creating instances of the vtdgaus class
setGeneric("vtdgaus", function(y, pstart = NULL, fixed = NULL, ...) standardGeneric("vtdgaus"))

# Define a method for creating instances of the vtdgaus class
setMethod("vtdgaus", 
          signature(y="ANY"), 
          function(y, yield, pstart=NULL, fixed=NULL, ...) {
            y <- matrix(y, length(y))
            x <- matrix(1)
            parameters <- list()
            npar <- 2
            if(is.null(fixed)) fixed <- as.logical(rep(0, npar))
            if(!is.null(pstart)) {
              if(length(pstart) != npar) stop("length of 'pstart' must be ", npar)
              parameters$mu <- pstart[1]
              parameters$sigma <- pstart[2]
            } else {
              parameters <- list(mu=.5, sigma=1)
            }
            mod <- new("vtdgaus", parameters=parameters, fixed=fixed, x=x, y=y, npar=npar, yield=yield)
            mod
          }
)

# Define method for displaying parameters of vtdgaus class
setMethod("show", "vtdgaus",
          function(object) {
            cat("Gaussian with variable discrete support for percentage responses\n")
            cat("Parameters: \n")
            cat("mu: ", object@parameters$mu, "\n")
            cat("sigma: ", object@parameters$sigma, "\n")
          }
)

# Define method for computing density of vtdgaus class
setMethod("dens", "vtdgaus",
          function(object, log=FALSE) {
            prec <- pnorm(1 + .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma) - 
                    pnorm(0 - .5*(1/60), mean = object@parameters$mu, sd = object@parameters$sigma)
            if(prec < 1e-12) {
              p <- 1/(object@yield + 1)
            } else {
              p <- pnorm(object@y + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - 
                   pnorm(object@y - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
              norm <- pnorm(1 + .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma) - 
                      pnorm(0 - .5*(1/object@yield), mean = object@parameters$mu, sd = object@parameters$sigma)
              p <- p/norm
            }
            p[object@yield == 0] <- 1
            if(log) return(log(p)) else return(p)
          }
)

# Define method for setting parameters of vtdgaus class
setMethod("setpars", "vtdgaus",
          function(object, values, which="pars", ...) {
            npar <- npar(object)
            if(length(values) != npar) stop("length of 'values' must be", npar)
            nms <- names(object@parameters)
            switch(which,
                   "pars"= {
                     object@parameters$mu <- values[1]
                     object@parameters$sigma <- values[2]
                   },
                   "fixed" = {
                     object@fixed <- as.logical(values)
                   }
            )
            names(object@parameters) <- nms
            return(object)
          }
)

# Define method for getting parameters of vtdgaus class
setMethod("getpars", "vtdgaus",
          function(object, which="pars", ...) {
            switch(which,
                   "pars" = {
                     parameters <- numeric()
                     parameters <- unlist(object@parameters)
                     pars <- parameters
                   },
                   "fixed" = {
                     pars <- object@fixed
                   }
            )
            return(pars)
          }
)

# Define method for fitting vtdgaus class
setMethod("fit", "vtdgaus",
          function(object, w) {
            if(missing(w)) w <- NULL
            if(!is.null(w)) {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(w*log(dens(object)))
              }
            } else {
              negLL <- function(pars) {
                object <- setpars(object, c(pars[1], exp(pars[2])))
                -sum(log(dens(object)))
              }
            }
            pars <- optim(c(object@parameters$mu, log(object@parameters$sigma)), fn=negLL)$par
            object <- setpars(object, c(pars[1], exp(pars[2])))
            object
          }
)

# Function for reordering states in an HMM based on labels
label_switch <- function(mod, labels) {
  if(!is(mod, "depmix") || !is(mod, "depmix.fitted")) stop("this function is for depmix models")
  n_states <- mod@nstates
  if(length(labels) != n_states || length(unique(labels)) != n_states || !(all(labels) %in% 1:n_states)) {
    stop("labels needs to be a vector of unique integers between 1 and", n_states)
  }
  inv_labels <- sapply(1:n_states, function(x) which(labels == x))
  tmp <- mod
  # Relabel prior
  ppars <- getpars(mod@prior)
  fpars <- getpars(mod@prior, which="fixed")
  out_pars <- as.numeric(t(matrix(ppars, nrow=length(ppars)/n_states, byrow=TRUE)[, inv_labels]))
  out_fixed <- as.logical(t(matrix(fpars, nrow=length(fpars)/n_states, byrow=TRUE)[, inv_labels]))
  if(!tmp@prior@family$link == "identity") tmp@prior@family$base <- labels[tmp@prior@family$base]
  # Relabel transition
  for(i in 1:n_states) {
    ppars <- getpars(mod@transition[[inv_labels[i]]])
    fpars <- getpars(mod@transition[[inv_labels[i]]], which="fixed")
    out_pars <- c(out_pars, as.numeric(t(matrix(ppars, nrow=length(ppars)/n_states, byrow=TRUE)[, inv_labels])))
    out_fixed <- c(out_fixed, as.logical(t(matrix(fpars, nrow=length(fpars)/n_states, byrow=TRUE)[, inv_labels])))
    tmp@transition[[i]] <- mod@transition[[inv_labels[i]]]
    if(!tmp@transition[[i]]@family$link == "identity") tmp@transition[[i]]@family$base <- labels[tmp@transition[[i]]@family$base]
  }
  # Relabel response
  for(i in 1:n_states) {
    out_pars <- c(out_pars, unlist(lapply(mod@response[[inv_labels[i]]], getpars)))
    out_fixed <- c(out_fixed, unlist(lapply(mod@response[[inv_labels[i]]], getpars, which="fixed")))
  }
  tmp <- setpars(tmp, out_fixed, which="fixed")
  tmp <- setpars(tmp, out_pars)
  if(is(tmp, "depmix.fitted")) tmp@posterior <- viterbi(tmp)
  return(tmp)
}

# Function for ordering states in a truncated discretised Gaussian model
order_mod_truncdiscgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0, ns)
  for(i in 1:ns) {
    # Calculate expected return in each state
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- truncdiscgaus(seq(0, 20), pstart=tpars, min=-0.5, max=20.5)
    sum[i] <- sum(seq(0, 20) * dens(dmod))
  }
  cat("Expected values for states: ", sum, "\n")
  # Reorder the states
  mod <- label_switch(mod, rank(sum, ties.method="first"))
  return(mod)
}

# Function for ordering states in a variable response truncated discretised Gaussian model
order_mod_vtdgaus <- function(mod) {
  ns <- nstates(mod)
  sum <- rep(0.0, ns)
  for(i in 1:ns) {
    # Calculate expected return in each state
    tpars <- getpars(mod@response[[i]][[1]])
    dmod <- vtdgaus(seq(0, 1, length=61), pstart=tpars, yield=rep(60, 61))
    sum[i] <- sum(seq(0, 1, length=61) * dens(dmod))
  }
  # Reorder the states
  mod <- label_switch(mod, rank(sum, ties.method="first"))
  return(mod)
}

```

```{r}

#| label: setup-parallel-3

# Set up the parallel backend
num_cores <- 8
#create the cluster
my.cluster <- parallel::makeCluster(
  num_cores,
  type = "FORK"
  )

#register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)
```

```{r}
#| label: fit-HMM-trustee-models
#| cache: true

# Depmix simple Models 
priordat <- data_HMM |>
  filter(roundNum==1)

tr_fdmod <- list()

max_ns <- 10

set.seed(20240513)
my_seeds <- sample(1000:50000, max_ns)

# Fitting HMM models to data 
tr_fdmod <- foreach(state = 2:max_ns, .packages = c("depmixS4"), .combine = "c") %dopar% {
  # Define depmix model with simple transition only depending on next_inv
  tr_simple <- depmix(returns ~ 1, data=data_HMM, nstates=state, transition=~ next_investment, prior=~ investment, initdata=priordat, family=multinomial("mlogit"), ntimes=rep(15,414))

  mus <- (1:state)*(1/(state+1))
  sigmas <- rep(1/(state+1), state)
  if(state > 8) sigmas <- sigmas + .1
  rModels_tr <- list()
  for(s in 1:state) {
    rModels_tr[[s]] <- list(vtdgaus(y=data_HMM$ret_pct_0, yield = 3*data_HMM$investment, pstart=c(mus[s],sigmas[s])))
  }

  tr_dmod <- makeDepmix(response=rModels_tr, transition=tr_simple@transition, prior=tr_simple@prior, ntimes = rep(15,414), homogeneous=FALSE)

  tmp1 <- fit(tr_dmod, emcontrol=em.control(random.start = FALSE))
  set.seed(my_seeds[state])
  tmp2 <- multistart(tr_dmod, nstart=20)
  if(!is.na(logLik(tmp2)) && logLik(tmp2) > logLik(tmp1)) {tmp1 <- tmp2}
  tmp1
}

# Stop the parallel backend
stopImplicitCluster()

save(tr_fdmod, file="data/tr_fdmod.RData", compression_level = 9)

```

```{r}
#|  label: HMM-trustee-model-comparison
#|  include: false

df_bics_tr <- data.frame(nstate = 2:max_ns,
                          logLik = unlist(lapply(tr_fdmod, logLik)),
                          unlist(lapply(tr_fdmod, function(x) attr(logLik(x), which="df"))),
                          AIC = unlist(lapply(tr_fdmod, AIC)),
                          BIC = unlist(lapply(tr_fdmod, BIC)))
df_bics_tr

# Here we choose BIC as best m
nstate_tr <- which.min(df_bics_tr[,5]) + 1

best_mod_tr_all <- order_mod_vtdgaus(tr_fdmod[[nstate_tr-1]])
summary(best_mod_tr_all)
```

```{r}
#| label: tbl-BICs-tr
#| warning: false
#| echo: false
#| tbl-cap: Model comparison of HMMs of trustee actions with different numbers of latent states.

df_bics_tr %>%
  mutate_if(is.numeric, format, digits=4, nsmall = 0, big.mark = ",") %>%
knitr::kable(booktabs = TRUE,
  col.names = c("#states", "log(lik)", "# par", "AIC", "BIC"))
```

```{r}
#| label: fig-trustee-policy-plot
#| echo: false
#| fig.cap: Probability mass function of the trustee's policy conditional on its latent state as an output of the best fitting XX state HMM using a truncated discretised Gaussian as a response function
#| fig.align: center
#| fig.width: 12
#| fig.height: 6

# Plotting the trustee HMM policy by state 
tr_policy <- list()
for(i in 1:nstates(best_mod_tr_all)) {
  tr_policy[[i]] <- dens(vtdgaus(seq(0,1, length=61), yield=3*20, pstart=c(unlist(best_mod_tr_all@response[[i]][[1]]@parameters))))  
}

perc_return <- seq(0,1, length=61)

respplot <- data.frame(state = rep(1:nstates(best_mod_tr_all), each=length(perc_return)),
                            perc_return = perc_return,
                            probability = unlist(tr_policy)) |>
  mutate(state = factor(paste0("s",state))) |> 
  ggplot(                 
       aes(x = perc_return,
           y = probability,
           fill = state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  # labs(fill='Latent trustee state') +
  xlab("Percentage return") +
  theme_minimal() + 
  theme(legend.position = "right") + ggtitle("State conditional strategies")

# Plotting Prior (initial) state probabilities
# priorplot <- data.frame(state=factor(paste0("s",1:nstate_tr)), prob=getpars(getmodel(best_mod_tr_all, "prior"))) |>
#   ggplot(aes(x = state, y = prob, fill=state)) +
#   geom_bar(stat = "identity", width = 0.5) +
#   xlab("state") +
#   ylab("prior probability") +
#   ggtitle("Initial state probability") +
#   theme_minimal()

# Function to calculate softmax probabilities
softmax <- function(x) exp(x) / sum(exp(x))

# Assuming best_mod_tr_all is your fitted model
params <- getpars(getmodel(best_mod_tr_all, "prior"))

# Extract intercepts and investment coefficients, and set initial investment value
intercepts <- params[1:nstate_tr]
investments <- params[(nstate_tr + 1):(2 * nstate_tr)]
investment_value <- 9

# Calculate logits and initial probabilities
logits <- intercepts + investments * investment_value
initial_probs <- softmax(logits)

# Create a data frame for plotting
priordf <- data.frame(
  state = factor(paste0("s", 1:nstate_tr)),
  prob = initial_probs
)

# Plot the initial state probabilities
priorplot <- ggplot(priordf, aes(x = state, y = prob, fill = state)) +
  geom_bar(stat = "identity", width = 0.5) +
  xlab("State") +
  ylab("Prior Probability") +
  ggtitle("Initial State Probability for average investment") +
  theme_minimal()

# Plotting Transition function between states
plot_transitions <- function(fmod,numCovIn,numCovTr) {

  ns <- nstates(fmod)

  trans_prob <- data.frame(
    from = rep(1:ns, each=80*ns),
    to = rep(1:ns, each=80),
    inv = seq(0,20,length=80),
    prob = 0
  )

  y <- matrix(0.0,ncol=ns, nrow=80)

  for(from in 1:ns) {
    #pars <- matrix(getpars(fmod)[seq(ns + 1 + (from - 1)*2*ns, ns + 1 + (from - 1)*2*ns + 2*ns - 1)],ncol=2)
    pars <- matrix(getpars(fmod)[seq(ns*numCovIn + 1 + (from - 1)*numCovTr*ns, ns*numCovIn + 1 + (from - 1)*numCovTr*ns + numCovTr*ns - 1)], ncol=numCovTr)
    paste0("from state: ", from)
    print(pars)
    for(to in 1:ns) {
      x <- trans_prob[trans_prob$from == from & trans_prob$to == to, "inv"]
      y[,to] <- exp(pars[to,1] + pars[to,2]*x)
    }
    y <- y/rowSums(y)
    for(to in 1:ns) {
      trans_prob$prob[trans_prob$from == from & trans_prob$to == to] <- y[,to]
    }
  }

  trans_prob <- trans_prob %>%
    mutate(from = paste0("s", from),
           to = paste0("s", to))

  ggplot(trans_prob,aes(x=inv,y=prob, colour = as.factor(to))) + geom_line() + facet_wrap(~from) + ylim(c(0,1))
}

transplot <- plot_transitions(best_mod_tr_all,numCovIn=2,numCovTr=2) + labs(x = "Investment", y = "Probability", color='Transition to') + theme_minimal() + theme(legend.position = "bottom")

gridExtra::grid.arrange(respplot, priorplot, transplot, ncol=3, layout_matrix=cbind(c(1,2),c(3,3),c(3,3)))

```